---
title: "Advisory Group Analysis"
author: "Janette Avelar"
date: '2023-05-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(here)
library(DT)
load(here("data", "complete_canopy_2023.Rdata"))
source(here("scripts", "branding.R"))
```

# What's the racial/ethnic breakdown of Canopy school teaching teams?
```{r teaching staff race}
teach_race <- confidential %>% 
  select(school_id, teaching_diversity) %>% 
  filter(!(teaching_diversity == 0)) %>% 
  filter(!(teaching_diversity == 5)) %>% 
  mutate(teaching_diversity = case_when(
    teaching_diversity == 1 ~ "0-24%\nBIPOC",
    teaching_diversity == 2 ~ "25-49%\nBIPOC",
    teaching_diversity == 3 ~ "50-74%\nBIPOC",
    teaching_diversity == 4 ~ "75-100%\nBIPOC"
  ))

ggplot(teach_race, aes(teaching_diversity, fill = teaching_diversity)) +
  geom_bar() +
  scale_fill_manual(values = transcend_cols) +
  theme(legend.position = "none")

teach_race_table <- teach_race %>% 
  mutate(rate = rep(1, length(244))) %>% 
  group_by(teaching_diversity) %>% 
  summarize(N = sum(rate),
            Pct = round(N/244, 2))
datatable(teach_race_table)
```

# What's the racial/ethnic breakdown of Canopy school leadership teams?
```{r leadership team race}
lead_race <- confidential %>% 
  select(school_id, leadership_diversity) %>% 
  filter(!(leadership_diversity == 0)) %>% 
  mutate(leadership_diversity = case_when(
    leadership_diversity == 1 ~ "0-24%\nBIPOC",
    leadership_diversity == 2 ~ "25-49%\nBIPOC",
    leadership_diversity == 3 ~ "50-74%\nBIPOC",
    leadership_diversity == 4 ~ "75-100%\nBIPOC"
  ))
#lead_race_plot <-
ggplot(lead_race, aes(leadership_diversity, fill = leadership_diversity)) +
  geom_bar() +
  scale_fill_manual(values = transcend_cols) +
  theme(legend.position = "none") +
  labs(title = "What is the racial/ethnic breakdown of Canopy school leadership teams?",
        x = "",
        y = "Number of Schools")
# ggsave("finding-4.png", plot = lead_race_plot, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")

lead_race_table <- lead_race %>% 
  mutate(rate = rep(1, length(251))) %>% 
  group_by(leadership_diversity) %>% 
  summarize(N = sum(rate),
            Pct = round(N/251, 2))
datatable(lead_race_table)
```

# How do these demographics compare to previous years?
Leadership diversity in our pool of Canopy schools increased this year. Last year, 28% of our schools were led by predominantly BIPOC leadership teams. This year, 41% of our schools are led by predominantly BIPOC leadership teams.

**Whoo hoo!**

# Did we meet our goal of oversampling BIPOC-led schools?
We may be oversampling in relation to national averages (I'm unsure where to find this) but we are not yet representing BIPOC-led schools as the majority of our sample. Still, increasing from 28% to 41% is a step in the right direction.

We hypothesized if we increased our nomination pool than we would see more BIPOC-led schools, given that our BIPOC-led nominating organizations and BIPOC-identifying nominators had historically nominated the majority of our BIPOC-led schools. We did succeed in diversifying our nomination pool, and this seems to have also diversified our school sample.