---
title: "canopy_school_size"
author: "Anwesha Guha"
date: "6/20/2023"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include = FALSE}
library(pacman)
p_load(here, tidyverse, psych, DT, outliers)
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```

## What was the average size of Canopy schools?

Average size is `r mean(full$self_reported_total_enrollment)` students. Let's visualize that below.

```{r fig.height=8, fig.width=12}
plot_size <- full %>% 
  ggplot(aes(self_reported_total_enrollment)) +
  geom_histogram(fill = transcend_cols[1], col = "black", bins = 45, linewidth = .2) +
  geom_vline(aes(xintercept = mean(self_reported_total_enrollment)), col = transcend_cols[3]) +
  geom_vline(aes(xintercept = median(self_reported_total_enrollment)), col = transcend_cols[3]) +  
  annotate("text", x=530, y=70, label="Mean = 450", size = 4, col = transcend_cols[3], angle = 270) +
  annotate("text", x=350, y=95, label="Median = 270", size = 4, col = transcend_cols[3], angle = 270) +
  labs(x = "Reported Enrollment",
       y = "Number of Schools", 
       title = "What was the average size of Canopy schools?") +
  ylim(0, 100)

plot_size

# ggsave("finding-3.png", plot = plot_size, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

Note that one really large outlier. The median is almost half that number: `r median(full$self_reported_total_enrollment)`

I modified the graph to include both, but you might choose to highlight one or the other.

## Janette modifications 6.21.23
Modified mean:

I replaced the extreme outlier value with the median value for our sample, and recalculated the average size. With imputation for that single value, the average size of Canopy schools is: 411

However, when creating thresholds for outliers using interquartile range, there are actually 10 outliers identified. With imputation for all 10 outliers, the average size of Canopy schools is: 333

```{r results = FALSE}
#check out quartiles
summary(full$self_reported_total_enrollment)
#get IQR
IQR(full$self_reported_total_enrollment)
#generate extreme threshold value 
#3rd quartile +/- (3*IQR)
tmin = 480-(3*354.5)
tmax = 480+(3*354.5)
#identify extreme outliers
full$self_reported_total_enrollment[which(full$self_reported_total_enrollment < tmin | full$self_reported_total_enrollment > tmax)]
#test if it's an extreme outlier
chisq.out.test(full$self_reported_total_enrollment)
#correcting for single extreme outlier
correct_avg1 <- full %>% 
  select(school_id, size = self_reported_total_enrollment) %>% 
  mutate(size = case_when(
    size == 10000 ~ median(size, na.rm = TRUE),
    TRUE ~ as.numeric(size)
  )) %>% 
  summarize(mean = mean(size))
#correcting for all outliers
correct_avg2 <- full %>% 
  select(school_id, size = self_reported_total_enrollment) %>% 
  mutate(size = case_when(
    size > tmax ~ median(size, na.rm = TRUE),
    TRUE ~ as.numeric(size)  
  )) %>% 
  summarize(mean = mean(size))
```

## What is the average size for Canopy high schools?

```{r}
full %>% 
  group_by(grades_high) %>% 
  summarize(mean = mean(self_reported_total_enrollment, na.rm = TRUE),
            median = median(self_reported_total_enrollment, na.rm = TRUE), 
            min = min(self_reported_total_enrollment),
            max = max(self_reported_total_enrollment))
  
```

## Are all outliers virtual schools?
No, only a few of the top schools are.
75% of schools are falling under ~500 students.

```{r}
outlier <- full %>% 
  select(school_name, school_descriptor_virtual, self_reported_total_enrollment, grades_elementary, grades_middle, grades_high, school_descriptor)
quantile(outlier$self_reported_total_enrollment, probs = c(0,0.25,0.5,0.75,1))
quantile(full$self_reported_total_enrollment, probs = c(0,0.25,0.5,0.75,1))
```

