---
title: 'Clustering Analysis Part 2: Breakdown'
author: "Anwesha Guha"
date: "5/22/2023"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)
p_load(here, purrr, tidyverse, ggcorrplot, proxy, plotly, patchwork, psych, GPArotation, parameters, DT, nFactors, ggthemes)
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```

# Parent/Child Breakdown

```{r}
parent <- c("practices_co_leaders", "practices_blended_learning", "practices_all_courses_designed_for_inclusion", "practices_design_marginalized")

child_leadership <- c("practices_teachers_leaders", "practices_students_leaders", "practices_families_leaders", "practices_community_leaders", "practices_industry_leaders", "practices_other_leaders")

child_blended <- c("practices_a_la_carte", "practices_flipped_classroom", "practices_flex", "practices_enriched_virtual", "practices_station_rotation")
```

```{r}
practices_cor <- tags %>%
  select(starts_with("practices")) %>%
  cor

efa_4 <- fa(practices_cor, nfactors = 4, rotate = "oblimin", fm = "minres")

# print(efa_4, sort = T)

efa_max <- efa_4 %>%
  model_parameters(sort = TRUE, threshold = "max") %>% 
  mutate(status = case_when(Variable == "practices_co_leaders" ~ "parent_co_leader",
                            Variable == "practices_blended_learning" ~ "parent_blended",
                            Variable == "practices_all_courses_designed_for_inclusion" ~ "parent_inclusion",
                            Variable == "practices_design_marginalized" ~ "parent_design_margin",
                         Variable %in% child_leadership ~ "child_leadership",
                         Variable %in% child_blended ~ "child_blended"))

efa_max %>%
  datatable(caption = "Max loadings") %>%
  formatRound(digits = 2, columns = 2:7) 

efa_4 %>%
  model_parameters(sort = TRUE, threshold = 0.28) %>%
  datatable(caption = "Threshold loadings") %>%
  formatRound(digits = 2, columns = 2:7) 

efa_4 %>%
  model_parameters(sort = TRUE) %>%
  datatable(caption = "All loadings") %>%
  formatRound(digits = 2, columns = 2:7) #this is the final EFA file
```

## EFA with child tags removed
```{r}
practices_cor_nonest <- tags %>%
  select(starts_with("practices")) %>%
  select(-c(all_of(child_leadership), all_of(child_blended))) %>% 
  cor
```

```{r}
efa_4_nonest <- fa(practices_cor, nfactors = 4, rotate = "oblimin", fm = "minres")

efa_4_nonest %>%
  model_parameters(sort = TRUE, threshold = "max") %>%
  datatable(caption = "Max loadings") %>%
  formatRound(digits = 2, columns = 2:7) 

child_c4 <- efa_4_nonest %>%
  model_parameters(sort = TRUE) # added 6/7 to look at code below for only child tags
```


# School Type Breakdown (Public, District, Charter)

```{r}
c4 <- read_tsv(here("output/EFA 2023 Results All.txt"))

cweights <- c4 %>% 
#  rename(all_of(clust_details_r)) %>%
  select(-Complexity, -Uniqueness) %>%
  pivot_longer(cols = -Variable, names_to = "cluster", values_to = "weight") %>%
  rename(practice = Variable)

practices_long <- tags %>% 
  select(school_id, starts_with("practices")) %>%
  pivot_longer(-school_id, names_to = "practice", values_to = "val")

clusters_by_school <- practices_long %>%
  left_join(cweights, by = "practice") %>%
  group_by(school_id, cluster) %>% 
  summarize(cluster_weight = sum(val * weight, na.rm = TRUE), .groups = "drop") %>%
  arrange(school_id)

clusters_by_school_wide <- clusters_by_school %>%
  pivot_wider(id_cols = school_id, names_from = cluster, values_from = cluster_weight) %>%
  arrange(school_id) %>% 
  rename(learning_mastery = MR1,
         holistic_student_justice = MR2,
         postsec_pathways_world = MR3,
         blended_flexible_learning = MR4)
```

## All tags
```{r echo=TRUE, fig.width=10, message=FALSE}
school_type_dat <- full %>%
  select(school_id, school_descriptor) %>%
  arrange(school_id) %>% 
  mutate(n = 1,
         school_descriptor = case_when(school_descriptor == 1 ~ "Public district",
                                       school_descriptor == 2 ~ "Public charter",
                                       school_descriptor == 3 ~ "Independent/private")) %>% 
  pivot_wider(names_from = school_descriptor,
              values_from = n, 
              values_fill = 0) %>% 
  select(-school_id)


school_type_clusters <- cor(school_type_dat, 
  clusters_by_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

ggcorrplot(school_type_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand school type",
         fill = "Correlation") +
    #scale_x_discrete(labels = label_tags) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```

## Child tags removed
```{r}
child_cweights <- child_c4 %>% 
#  rename(all_of(clust_details_r)) %>%
  select(-Complexity, -Uniqueness) %>%
  pivot_longer(cols = -Variable, names_to = "cluster", values_to = "weight") %>%
  rename(practice = Variable)

child_practices_long <- tags %>% 
  select(school_id, starts_with("practices")) %>%
  select(-c(all_of(child_leadership), all_of(child_blended))) %>% 
  pivot_longer(-school_id, names_to = "practice", values_to = "val")

child_clusters_by_school <- child_practices_long %>%
  left_join(child_cweights, by = "practice") %>%
  group_by(school_id, cluster) %>% 
  summarize(cluster_weight = sum(val * weight, na.rm = TRUE), .groups = "drop") %>%
  arrange(school_id)

child_clusters_by_school_wide <- child_clusters_by_school %>%
  pivot_wider(id_cols = school_id, names_from = cluster, values_from = cluster_weight) %>%
  arrange(school_id) %>% 
  rename(learning_mastery = MR1,
         holistic_student_justice = MR2,
         postsec_pathways_world = MR3,
         blended_flexible_learning = MR4)
```

```{r}
child_school_type_clusters <- cor(school_type_dat, 
  child_clusters_by_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

ggcorrplot(child_school_type_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand school type (child tags omitted)",
         fill = "Correlation") +
    #scale_x_discrete(labels = label_tags) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```


# School Locale Breakdown (Rural, Suburban, Urban)

## All tags

```{r echo=TRUE, fig.width=10, message=FALSE}
school_locale_dat <- full %>%
  select(school_id, self_reported_locale_rural, self_reported_locale_urban, self_reported_locale_suburban) %>%
  arrange(school_id) %>% 
  select(-school_id)


school_locale_clusters <- cor(school_locale_dat, 
  clusters_by_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

ggcorrplot(school_locale_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand school locale",
         fill = "Correlation") +
    scale_x_discrete(labels = c("Rural", "Urban", "Suburban")) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```

## Child tags removed
```{r}
child_school_locale_clusters <- cor(school_locale_dat, 
  child_clusters_by_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

ggcorrplot(child_school_locale_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand locale (child tags omitted)",
         fill = "Correlation") +
    #scale_x_discrete(labels = label_tags) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```

# Locale Breakdown by HS Status

Filter by high school (and not high school) and look at correlations with locale.

```{r}
# Filter schools by type
hs_dat <- full %>% 
  select(school_id, grades_high) %>% 
  filter(grades_high == 1)

not_hs_dat <- full %>% 
  select(school_id, grades_high) %>% 
  filter(grades_high == 0)

# Filter schools by type in tags
hs_tags <- left_join(hs_dat, tags) %>% 
  select(-grades_high)

not_hs_tags <- left_join(not_hs_dat, tags) %>% 
  select(-grades_high)

# Create correlation matrices
hs_practices_cor <- hs_tags %>%
  select(starts_with("practices")) %>%
  cor

not_hs_practices_cor <- not_hs_tags %>%
  select(starts_with("practices")) %>%
  cor

# Run EFA
hs_efa_4 <- fa(hs_practices_cor, nfactors = 4, rotate = "oblimin", fm = "minres") %>% 
  model_parameters(sort = TRUE)

not_hs_efa_4 <- fa(not_hs_practices_cor, nfactors = 4, rotate = "oblimin", fm = "minres") %>% 
  model_parameters(sort = TRUE)

```

High school corr
```{r}

hs_cweights <- hs_efa_4 %>% 
#  rename(all_of(clust_details_r)) %>%
  select(-Complexity, -Uniqueness) %>%
  pivot_longer(cols = -Variable, names_to = "cluster", values_to = "weight") %>%
  rename(practice = Variable)

hs_practices_long <- hs_tags %>% 
  select(school_id, starts_with("practices")) %>%
  pivot_longer(-school_id, names_to = "practice", values_to = "val")

clusters_by_high_school <- hs_practices_long %>%
  left_join(hs_cweights, by = "practice") %>%
  group_by(school_id, cluster) %>% 
  summarize(cluster_weight = sum(val * weight, na.rm = TRUE), .groups = "drop") %>%
  arrange(school_id)

clusters_by_high_school_wide <- clusters_by_high_school %>%
  pivot_wider(id_cols = school_id, names_from = cluster, values_from = cluster_weight) %>%
  arrange(school_id) %>% 
  rename(learning_mastery = MR1,
         holistic_student_justice = MR2,
         postsec_pathways_world = MR3,
         blended_flexible_learning = MR4)
```

```{r}
high_school_locale_dat <- full %>%
  filter(grades_high == 1) %>% 
  select(school_id, self_reported_locale_rural, self_reported_locale_urban, self_reported_locale_suburban) %>%
  arrange(school_id) %>% 
  select(-school_id)

not_high_school_locale_dat <- full %>%
  filter(grades_high == 0) %>% 
  select(school_id, self_reported_locale_rural, self_reported_locale_urban, self_reported_locale_suburban) %>%
  arrange(school_id) %>% 
  select(-school_id)


high_school_locale_clusters <- cor(high_school_locale_dat, 
  clusters_by_high_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

not_high_school_locale_clusters <- cor(high_school_locale_dat, 
  clusters_by_high_school_wide %>% 
  select(-school_id),
  method = "spearman"
)

ggcorrplot(high_school_locale_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand school locale for high schools",
         fill = "Correlation") +
    scale_x_discrete(labels = c("Rural", "Urban", "Suburban")) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```

```{r}
ggcorrplot(not_high_school_locale_clusters, hc.order = FALSE, lab = TRUE, lab_size = 2) +
    scale_fill_distiller(type = "div", limits = c(-1, 1), expand = c(0, 0)) +
    labs(title = "Correlation between practice clusters\nand school locale for NOT high schools",
         fill = "Correlation") +
    scale_x_discrete(labels = c("Rural", "Urban", "Suburban")) +
    #scale_y_discrete(labels = label_tags) +
    labs(x = "", y = "") + 
    #scale_fill_cc_gradient + 
    theme_transcend_sparse + 
#    scale_y_discrete(labels = label_clust) +
    theme(
      axis.text.x = element_text(angle = 35, hjust = 1, vjust = 1, size = rel(0.8)),
      axis.text.y = element_text(size = rel(0.8)),
      panel.border = element_blank(),
      axis.ticks = element_blank()
    )
```


# School Demographic Profile Breakdown (%BIPOC students and/or staff)
