---
title: "final_modeling_for_findings"
author: "Anwesha Guha"
date: "6/23/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(here, purrr, tidyverse, ggcorrplot, proxy, plotly, patchwork, psych, GPArotation, parameters, DT, nFactors, broom, scales, rstanarm, broom.mixed, gmodels)

source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
load(here("data/demo_models_for_findings.RData"))
```

# DONT RUN ANY OF THE BELOW -- SAVED AUTOMATICALLY IN .RDATA & WILL BE TOO TIME CONSUMING

Cleaning
```{r data prep}
mod_prep <- full %>% 
  select(school_id, school_descriptor, starts_with("practices_"), starts_with("self_reported"), starts_with("grades_")) %>% 
  select(-c(starts_with("self_reported_grades"), self_reported_locale_multiple, self_reported_ela, self_reported_math, starts_with("self_reported_locale_other"), ends_with("none_collected"))) %>% 
  pivot_longer(cols = c(starts_with("practices")),
               names_to = "tag",
               values_to = "value")

mod_prep <- mod_prep %>% 
  mutate(black_percent = self_reported_race_black/self_reported_total_enrollment, 
         hispanic_percent = self_reported_race_hispanic/self_reported_total_enrollment,
         bipoc_percent = 1 - (self_reported_race_white/self_reported_total_enrollment),
         native_percent = self_reported_race_aian/self_reported_total_enrollment,
         asian_percent = self_reported_race_asian/self_reported_total_enrollment,
         frpl_percent = self_reported_frpl/self_reported_total_enrollment,
         swd_percent = self_reported_swd,
         ell_percent = self_reported_ell,
         school_descriptor = case_when(school_descriptor == 1 ~ "public_district",
                                       school_descriptor == 2 ~ "public_charter",
                                       school_descriptor == 3 ~ "independent_private")) %>% 
  rename(locale_urban = self_reported_locale_urban,
         locale_rural = self_reported_locale_rural,
         locale_suburban = self_reported_locale_suburban)

#pivot school type/descriptor variable for more consistent analysis, like locale and level
library(fastDummies)
mod_prep <- mod_prep %>% 
  dummy_cols(select_columns = "school_descriptor")

#get rid of the columns I don't want (not recommended for replication)
mod_prep <- mod_prep %>% 
  select(-c(3:4,8:15))


## below is filtering on tags that have been selected at least 25 times
m <- mod_prep %>% 
  group_by(tag, value, school_id) %>% 
  summarise(selected_n = n()) %>% 
  arrange(selected_n) %>% 
  filter(value > 0) %>% 
  select(-2)

s <- full %>% 
  group_by(school_id) %>% 
  summarise(count = n())

## we see that low select tags are the following
# modified to those selected less than 25%
low_select_tags <- c("practices_other_leaders", "practices_heritage_language", "practices_enriched_virtual",
                     "practices_translanguaging", "practices_flipped_classroom", "practices_a_la_carte", "practices_dual_language", "practices_industry_leaders", "practices_assessments_bilingual", "practices_flex", "practices_community_leaders", "practices_station_rotation", "practices_assessments_agency", "practices_families_leaders", "practices_interoperability")

## remove them for our analysis
mod_prep <- mod_prep %>% 
  filter(!(tag %in% low_select_tags))


one_vars = c("self_reported_total_enrollment", "black_percent", "hispanic_percent", "bipoc_percent", "native_percent", "asian_percent", "locale_urban", "locale_rural", "locale_suburban", "frpl_percent", "swd_percent", "ell_percent", "school_descriptor_independent_private", "school_descriptor_public_charter", "school_descriptor_public_district","grades_prek", "grades_elementary", "grades_middle", "grades_high")


logistic_one_dat = 
  mod_prep %>%
  select(value, tag, one_of(one_vars)) %>%
  na.omit() %>%
  group_by(tag) %>%
  # scaling non-binary predictors
  mutate_at(vars(
    c(
      "self_reported_total_enrollment",
      "black_percent",
      "hispanic_percent",
      "bipoc_percent",
      "native_percent",
      "asian_percent",
      "frpl_percent",
      "swd_percent",
      "ell_percent"
    )
  ), scale)
```

Going to run these separately to make combining/modeling/formatting a bit easier. Again, clunky code and not recommended for replication.
```{r}

# [tag] <- locale_rural ~ self_reported_total_enrollment + black_percent + hispanic_percent + white_percent + native_percent + asian_percent + frpl_percent + swd_percent + ell_percent + school_descriptor + grades_prek + grades_elementary + grades_middle + grades_high

one_form = as.formula(paste("value", "~", paste(one_vars, collapse = "+")))

#plot(xm, "areas", prob = 0.9, prob_outer = 1)

bayes_mods = list()


for(x in unique(mod_prep$tag)) {
  bayes_mods[[x]] = stan_glm(
    one_form,
    data = filter(logistic_one_dat, tag == x),
    family = binomial(link = "logit"),
    prior = student_t( #df and scale updated to have less extreme values
      df = 10,
      location = 0,
      scale = 2.5
    )
  )
  
}

bayes_tidy = lapply(bayes_mods, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") %>%
  mutate(
    nice_tag = label_tags()(response),
    nice_demog = factor(label_dems(term))
  )
```

Make plots
```{r}
bayes_facet_dat = bayes_tidy %>% 
  group_by(term) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10) %>% 
  left_join(tag_labels, by = c("response" = "tag")) #rerun this chunk (not in saved r data)

list_plot <- bayes_facet_dat %>% 
  split(.$term) %>% 
  map(~ggplot(., aes(y = exp(estimate), 
             x = reorder_within(label, estimate, within = term, fun = mean)
             )) +
  geom_col(fill = transcend_cols[1]) + 
  labs(y = "",
       x = "") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .25, .5, 1, 2, 4, 16),
    labels = c("1/16 as likely", "1/4 as likely", "1/2 as likely", "Even", "2x as likely", "4x as likely", "16x as likely"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  geom_hline(yintercept = 1, color = transcend_grays[1]) +
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = rel(1)), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -45, size = rel(0.5), vjust = 0.5, hjust = 0))
  )
```

```{r}
file_names <- stringr::str_c(names(list_plot), ".png")
```

Run code below only to resave files
```{r}
pwalk(list(file_names, list_plot), ggsave, path = "final_products/draft-findings/finding-9")
```

```{r}
list_plot
```


# CAN START RUNNING CODE FROM BELOW ONWARDS 

Plots grouped to level
```{r}
prek_plot <- list_plot$grades_prek +
  geom_col(fill = transcend_cols[1]) +
  labs(title = "Pre-K Schools")

elem_plot <- list_plot$grades_elementary +
  geom_col(fill = transcend_cols[2]) +
  labs(title = "Elementary Schools")

middle_plot <- list_plot$grades_middle +
  geom_col(fill = transcend_cols[3]) +
  labs(title = "Middle Schools")

high_plot <- list_plot$grades_high +
  geom_col(fill = transcend_cols[4]) +
  labs(title = "High Schools")
```

```{r}
library(cowplot)
level_plot <- plot_grid(prek_plot, elem_plot, middle_plot, high_plot)
level_title <- ggdraw() + draw_label("Top 10 Practices That are Most/Least Likely to be Selected by Each School Level \nNote: Practices used by less than 25% of schools are omitted.", 
                                      fontface = "bold",
                                      color = transcend_grays[2],
                                      fontfamily = "Bebas Neue")

level_practices_plot <- plot_grid(level_title, level_plot, ncol = 1, rel_heights = c(0.1, 1))

ggsave("finding-9a.png", plot = level_practices_plot, path = here("final_products", "draft-findings"), width = 14, height = 8, units = "in")
```

Plots grouped to locale
```{r}
rural_plot <- list_plot$locale_rural +
  geom_col(fill = transcend_cols[1]) +
  labs(title = "Rural Schools")

urban_plot <- list_plot$locale_urban +
  geom_col(fill = transcend_cols[2]) +
  labs(title = "Urban Schools")

suburban_plot <- list_plot$locale_suburban +
  geom_col(fill = transcend_cols[3]) +
  labs(title = "Suburban Schools")
```

```{r}
locale_plot <- plot_grid(rural_plot, urban_plot, suburban_plot, align = "hv", nrow = 3)
locale_title <- ggdraw() + draw_label("Top 10 Practices That are Most/Least Likely to be Selected by Each Locale \nNote: Practices used by less than 25% of schools are omitted.", 
                                      fontface = "bold",
                                      color = transcend_grays[2],
                                      fontfamily = "Bebas Neue")

locale_practices_plot <- plot_grid(locale_title, locale_plot, ncol = 1, rel_heights = c(0.1, 1))

ggsave("finding-9b.png", plot = locale_practices_plot, path = here("final_products", "draft-findings"), width = 10, height = 12, units = "in")
```

Plots grouped to school descriptor
```{r}
independent_plot <- list_plot$school_descriptor_independent_private +
  geom_col(fill = transcend_cols[1]) +
  labs(title = "Independent Schools")

charter_plot <- list_plot$school_descriptor_public_charter +
  geom_col(fill = transcend_cols[2]) +
  labs(title = "Charter Schools")

district_plot <- list_plot$school_descriptor_public_district +
  geom_col(fill = transcend_cols[3]) +
  labs(title = "District Schools")
```

```{r}
schooltype_plot <- plot_grid(independent_plot, charter_plot, district_plot, align = "hv", nrow = 3)
schooltype_title <- ggdraw() + draw_label("Top 10 Practices That are Most/Least Likely to be Selected by Each School Type \nNote: Practices used by less than 25% of schools are omitted.", 
                                      fontface = "bold",
                                      color = transcend_grays[2],
                                      fontfamily = "Bebas Neue")

schooltyple_practices_plot <- plot_grid(schooltype_title, schooltype_plot, ncol = 1, rel_heights = c(0.1, 1))

ggsave("finding-9c.png", plot = schooltyple_practices_plot, path = here("final_products", "draft-findings"), width = 10, height = 12, units = "in")
```

Plots grouped to demographic subgroups
```{r}
bipoc_plot <- list_plot$bipoc_percent +
  geom_col(fill = transcend_cols[1]) +
  labs(title = "Schools with higher-than-average \n% students classified as \nBIPOC")

swd_plot <- list_plot$swd_percent +
  geom_col(fill = transcend_cols[2]) +
  labs(title = "Schools with higher-than-average \n% students classified as \nwith disabilities")

frpl_plot <- list_plot$frpl_percent +
  geom_col(fill = transcend_cols[3]) +
  labs(title = "Schools with higher-than-average \n% students eligible for \nfree/reduced price lunch")

ell_plot <- list_plot$ell_percent +
  geom_col(fill = transcend_cols[4]) +
  labs(title = "Schools with higher-than-average \n% students classified as \nEnglish language learners")
```

```{r}
library(cowplot)
demo_plot <- plot_grid(bipoc_plot, ell_plot, frpl_plot, swd_plot, align = "hv")
demo_title <- ggdraw() + draw_label("Top 10 Practices That are Most/Least Likely to be Selected by Student Demographic Characteristics \nNote: Practices used by less than 25% of schools are omitted.", 
                                      fontface = "bold",
                                      color = transcend_grays[2],
                                      fontfamily = "Bebas Neue")

demo_practices_plot <- plot_grid(demo_title, demo_plot, ncol = 1, rel_heights = c(0.1, 1))

ggsave("finding-9d.png", plot = demo_practices_plot, path = here("final_products", "draft-findings"), width = 18, height = 10, units = "in")
```
