---
title: "Misc. Follow-ups"
author: "Anwesha Guha"
date: "7/16/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
pacman::p_load(rio, here, tidyverse, ggrepel, DT, forcats)
```

```{r}
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```

## Guiding Questions

A) How many charter and district schools (so, excluding independent) are NEITHER part of a system-wide initiative NOR using a "model" in any way? (We had a survey question about "are you part of a system-wide innovation initiative", and another question about "to what degree are you using any of the following adoptable models." I think you'd want to look at schools that said "no" to the system-wide initiative, AND also said "not at all" to all of the 6 model options - Big Picture Learning, EL Education, etc.)


B) Basically the reverse - how many charter and district schools are BOTH part of a system-wide initiative, AND using at least 1 model ("somewhat" or "to a large extent" would both count as yes)

## Data prep

```{r}
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```

```{r}
partial <- full %>% 
  select(school_id, 
         school_name,
         school_state,
         school_descriptor,
         innovation_initiative,
         "Big Picture \nLearning"=model_usage_bpl,
         "EL Edu"=   model_usage_ele ,
         "IB" = model_usage_ib,
         "New Tech \nNetwork" = model_usage_ntn,
         "Opportunity \nCulture" =  model_usage_oc,
         "Summit \nLearning" = model_usage_sl)
```


Schools reported implementing adoptable models:
```{r}
#Check whether or not school implemented any adoptable models
partial %>% 
  mutate(across(c(6:11),
                ~ case_when(. > 0 ~ "Yes",TRUE ~ "No"))) %>% 
  pivot_longer(cols = 6:11, 
               names_to = "models",
               values_to = "response") %>% 
  ggplot(aes(x=models, 
             fill=response))+
  geom_bar(stat = "count", position = "dodge") + 
   scale_fill_manual(name="",
   values = transcend_cols) + 
  geom_text(stat = "count", aes(label = after_stat(count)), position = position_dodge(width = 1), vjust = -1, size = 3) + 
  labs(title = "How many schools reported implementing adoptable models?",
       x="")

```

```{r}
partial %>% 
mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 6:11, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0"))) %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..])),
            position = position_fill(vjust = 0.5),
            stat = "count",
            size = 3,
            color = "white")+
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "To what degree is your school utilizing any of the adoptable models listed below?",
       ) +
  theme(legend.position = "bottom") 
```

Charter vs district
```{r fig.height=10, fig.width=18}
partial %>%
  mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 6:11, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0")),
         school_descriptor = recode(school_descriptor, 
                              "1" = "Public district",
                              "2" = "Public Charter",
                              "3" = "Independent/private")) %>% 
  filter(school_descriptor != "Independent/private") %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="white", size =3)+
  scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "To what degree is your school utilizing any of the adoptable models listed below?",
       ) +
  facet_wrap(~school_descriptor) + 
  theme(legend.position = "bottom",
        axis.text = element_text(size = rel(1.2)),
        title = element_text(size = rel(1.6))) 
  
```

Part of a larger innovation initiative?
```{r fig.height=10, fig.width=18}
partial %>%
  mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 6:11, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0")),
         school_descriptor = recode(school_descriptor, 
                              "1" = "Public district",
                              "2" = "Public Charter",
                              "3" = "Independent/private")) %>% 
  filter(school_descriptor != "Independent/private") %>% 
  filter(innovation_initiative == 0) %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="white", size =3)+
  scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "For schools NOT a part of a systems-based initiative, \nto what degree are schools utilizing any of the adoptable models listed below?",
       ) +
  facet_wrap(~school_descriptor) + 
  theme(legend.position = "bottom",
        axis.title = element_text(size = rel(1.5)),
        axis.title.x = element_blank(),
        strip.text = element_text(size = rel(1.2)),
        legend.box.background = element_rect(fill="white", color="black"),
        title = element_text(size = rel(1.6))) 
```

```{r fig.height=10, fig.width=18}
partial %>%
  mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 6:11, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0")),
         school_descriptor = recode(school_descriptor, 
                              "1" = "Public district",
                              "2" = "Public Charter",
                              "3" = "Independent/private")) %>% 
  filter(school_descriptor != "Independent/private") %>% 
  filter(innovation_initiative == 1) %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="white", size =3)+
  scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "For schools that are a part of a systems-based initiative, \nto what degree are schools utilizing any of the adoptable models listed below?",
       ) +
  facet_wrap(~school_descriptor) + 
  theme(legend.position = "bottom",
        axis.title = element_text(size = rel(1.5)),
        axis.title.x = element_blank(),
        strip.text = element_text(size = rel(1.2)),
        legend.box.background = element_rect(fill="white", color="black"),
        title = element_text(size = rel(1.6))) 
```

# Tasks for cover essay

## Figure 12 -- assessment frequency bar
```{r}
assessments_dat <- full %>% 
  select(school_id, practices_assessments_agency, practices_assessments_deeper, practices_assessments_career, practices_assessments_sel)

assessments_dat <- assessments_dat %>% 
  pivot_longer(-school_id,
               names_to = "Assessment Type",
               values_to = "N")

assessments_dat2 <- assessments_dat %>% 
  group_by(`Assessment Type`) %>% 
  summarise(total = sum(N)) %>% 
  mutate(pct = total/251)
```

```{r}
assessments_plot <-
ggplot(assessments_dat2, aes(`Assessment Type`, pct, fill = `Assessment Type`)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Non-traditional outcomes that Canopy schools reported assessing",
       subtitle = "Note: Some schools selected multiple assessment types.\nTotal Canopy learning environments = 251",
       x = "",
       y = "Percentage of Canopy schools") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("Agency & \nSelf-Direction", "Career \nReadiness", "Deeper \nLearning", "Social-Emotional \nSkills")) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
```

```{r}
ggsave_transcend(assessments_plot, "cover_figure12", dir = here("final_products/essay-charts/cover_figures"), fig_width = 12, fig_height = 8)
```

## Figure 13 -- source of assessment
```{r}
assessment_source <- full %>% 
  select(school_id, starts_with("partnership_materials_"), starts_with("vendor_materials_"), starts_with("created_materials_"))

#remove bilingual assessment vars
assessment_source <- assessment_source %>% 
  select(-contains("bilingual"))

#create column that will help multiple be selected

assessment_source$partnership_materials_sum <- rowSums(assessment_source[ , c(2:6)], na.rm=TRUE)
assessment_source$vendor_materials_sum <- rowSums(assessment_source[ , c(7:11)], na.rm=TRUE)
assessment_source$created_materials_sum <- rowSums(assessment_source[ , c(12:16)], na.rm=TRUE)


#simplified materials vars
assessment_source <- assessment_source %>% 
  mutate(`combination of 2 or more` = case_when(
    (partnership_materials_sum != 0 & vendor_materials_sum != 0) ~ 1,
    (partnership_materials_sum != 0 & created_materials_sum != 0) ~ 1,
    (vendor_materials_sum != 0 & created_materials_sum != 0) ~ 1,
    TRUE ~ 0
         ),
    `partner-created assessments` = ifelse(partnership_materials_sum > 0, 1, 0),
    `vendor-created assessments` = ifelse(vendor_materials_sum > 0, 1, 0),
    `self-created assessments` = ifelse(created_materials_sum > 0, 1, 0))
```

```{r}
assessment_viz_dat <- assessment_source %>% 
  select(1, 20:23) %>% 
  pivot_longer(-school_id,
               names_to = "Assessment",
               values_to = "Times Selected")

assessment_viz_dat2 <- assessment_viz_dat %>% 
  group_by(`Assessment`) %>% 
  summarise(total = sum(`Times Selected`)) %>% 
  mutate(pct = total/251)

```

With bilingual assessments in the totals:
```{r}
assessment_source <- full %>% 
  select(school_id, starts_with("partnership_materials_"), starts_with("vendor_materials_"), starts_with("created_materials_"))

#create column that will help multiple be selected

assessment_source$partnership_materials_sum <- rowSums(assessment_source[ , c(2:7)], na.rm=TRUE)
assessment_source$vendor_materials_sum <- rowSums(assessment_source[ , c(8:13)], na.rm=TRUE)
assessment_source$created_materials_sum <- rowSums(assessment_source[ , c(14:19)], na.rm=TRUE)


#simplified materials vars
assessment_source <- assessment_source %>% 
  mutate(
    `partner-created assessments` = ifelse(partnership_materials_sum > 0, 1, 0),
    `vendor-created assessments` = ifelse(vendor_materials_sum > 0, 1, 0),
    `self-created assessments` = ifelse(created_materials_sum > 0, 1, 0),
    `combination of 2 or more` = case_when(
    (`partner-created assessments` + `vendor-created assessments` + `self-created assessments` > 1) ~ 1,
    TRUE ~ 0)
    )

assessment_viz_dat <- assessment_source %>% 
  select(1, 23:26) %>% 
  pivot_longer(-school_id,
               names_to = "Assessment",
               values_to = "Times Selected")

assessment_viz_dat2 <- assessment_viz_dat %>% 
  group_by(`Assessment`) %>% 
  summarise(total = sum(`Times Selected`)) %>% 
  mutate(pct = total/251)

```


```{r}
assessments_source_plot <-
ggplot(assessment_viz_dat2, aes(`Assessment`, pct, fill = `Assessment`)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Source of Assessments for Non-Traditional Outcomes",
       subtitle = "Note: Some schools selected multiple sources for assessments. \nTotal Canopy learning environments = 251",
       x = "",
       y = "") +
  theme(legend.position = "none") +
 scale_x_discrete(labels = c("Combination \nof 2 or more", "Partner-Created \nAssessment", "Self-Created \nAssessment", "Vendor-Created \nAssessment")) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
```

```{r}
ggsave_transcend(assessments_source_plot, "cover_figure13", dir = here("final_products/essay-charts/cover_figures"), fig_width = 12, fig_height = 8)
```

#Models--Innovation was not working for me for some reason, so code for systems_essay are below.

```{r}
model <- full %>% 
  select(school_id, 
         school_name,
         school_state,
         "Big Picture \nLearning"=model_usage_bpl,
         "EL Edu"=   model_usage_ele ,
         "IB" = model_usage_ib,
         "New Tech \nNetwork" = model_usage_ntn,
         "Opportunity \nCulture" =  model_usage_oc,
         "Summit \nLearning" = model_usage_sl,
         grades_high,
         self_reported_locale_rural,
         self_reported_locale_suburban,
         self_reported_locale_urban,
         self_reported_locale_multiple,
         waiver_application,
         political_impact)
```

## Anwesha edit
```{r}
adopt_mods_dat <- model %>% 
  mutate(across(c(4:9),
                ~ case_when(. > 0 ~ "Yes", TRUE ~ "No"))) %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  select(school_id, models, response) %>% 
  filter(response == "Yes")
```

```{r}
adopt_mods_dat <- adopt_mods_dat %>% 
  group_by(models) %>% 
  summarise(pct = round(length(school_id)/251*100)) 

adopted_mod_viz <- adopt_mods_dat %>% 
  ggplot(aes(x = fct_reorder(models, pct), y = pct)) +
  geom_col(fill = "#1A4C81") +
#  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(
    limits = c(0, 100),
    labels = unit_format(unit = "%")
  #  expand = expansion(mult = c(0, 0))
  ) +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Rate of adoption of six “models” by Canopy schools",
       subtitle = "Percentages show the proportion of all Canopy schools that indicated they are using each model “somewhat” or \n“to a large extent.” The remainder selected “Not at all” or did not respond to the question.",
       x = "",
       y = "Percentage of all Canopy schools ") +
  theme(legend.position = "none") +
# scale_x_discrete(labels = c("Combination \nof 2 or more", "Partner-Created \nAssessment", "Self-Created \nAssessment", "Vendor-Created \nAssessment")) +
  geom_text(aes(label = paste0(pct, "%")), 
            nudge_y = .02,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
```

```{r}
ggsave_transcend(adopted_mod_viz, "systems_figure1", dir = here("final_products/essay-charts/systems_figures"), fig_width = 12, fig_height = 8)
```


