---
title: "School State and Descriptors "
author: "Karen Zyskind"
date: "2023-04-28"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(reshape2)
library(cowplot)
library(dplyr)
library("pacman")
library("DT")
pacman::p_load("summarytools")
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2023.RData"))
```
**Questions about the data**   
I loaded what I think is the branding, but I am not sure if it worked or what I should see. 

##  Question 1 How many schools per state:
Table shows that the state with the most schools is California (21), followed
by Colorado (18), Massachusetts (15), and New York also with (15). 

```{r state_counts, echo = FALSE}
library(dplyr)

state_counts <- schools %>%
  group_by(school_state) %>%
  summarize(N_schools = n())

state_counts

# Load the knitr package
library(knitr)

# Created a table of the number of schools per state
n_schools <- aggregate(schools$school_name, 
                       by=list(State=schools$school_state), 
                       FUN=length)

# Renamed the columns of the new data frame
names(n_schools) <- c("State", "N_schools")

# Created a table using kable
kable(n_schools, format = "markdown", caption = "Number of schools per state")


```

## Question2 What is the distribution of descriptors by state?
Arizona, there were 3 schools with the descriptor "hybrid", 2 schools with the descriptor "microschool", 3 schools with the descriptor "other", and so on.

 Massachusetts had the most schools with the "other_text" descriptor (5), while Colorado had the most schools with the "microschool" descriptor (5).


```{r descriptor_count, results='asis', echo=FALSE}
library(dplyr)
library(DT)

descriptor_count <- schools %>% 
  group_by(school_state) %>% 
  summarise(homeschool = sum(school_descriptor_homeschool),
            hybrid = sum(school_descriptor_hybrid),
            microschool = sum(school_descriptor_microschool),
            other = sum(school_descriptor_other),
            other_text = sum(ifelse(is.na(school_descriptor_other_text), 0, nchar(school_descriptor_other_text) > 0)),
            sws = sum(school_descriptor_sws),
            virtual = sum(school_descriptor_virtual))

datatable(descriptor_count, 
          options = list(pageLength = 10),
          caption = "Descriptor Counts by State")
```


```{r}
#OR
# Created a new data frame with the counts of each descriptor by state
descriptor_count <- schools %>% 
  group_by(school_state) %>% 
  summarise(homeschool = sum(school_descriptor_homeschool),
            hybrid = sum(school_descriptor_hybrid),
            microschool = sum(school_descriptor_microschool),
            other = sum(school_descriptor_other),
           other_text = sum(ifelse(is.na(school_descriptor_other_text), 0, nchar(school_descriptor_other_text) > 0)),
            sws = sum(school_descriptor_sws),
            virtual = sum(school_descriptor_virtual))

# View the new data frame
descriptor_count

#to see the the description for other text since some schools wrote in their responses 
schools %>% 
  group_by(school_state, school_descriptor_other_text) %>% 
  summarise(n = n()) %>% 
  arrange(school_state, desc(n))




```

## Question 3 What is the breakdown of the schools by descriptor? 
```{r}

library(dplyr)
library(DT)

descriptor_count <- schools %>%
  select(school_descriptor_homeschool, school_descriptor_hybrid, school_descriptor_microschool, school_descriptor_other, school_descriptor_other_text, school_descriptor_sws, school_descriptor_virtual) %>%
  summarize(homeschool = sum(school_descriptor_homeschool),
            hybrid = sum(school_descriptor_hybrid),
            microschool = sum(school_descriptor_microschool),
            other = sum(school_descriptor_other),
            other_text = sum(ifelse(!is.na(school_descriptor_other_text), 1, 0)),
            sws = sum(school_descriptor_sws),
            virtual = sum(school_descriptor_virtual))

datatable(descriptor_count, caption = "Counts of School Descriptors")






```
## ##How do descriptors match up with school type?
```{r}
library(dplyr)
library(tidyr)
library(DT)

type <- schools %>%
  select(school_descriptor, school_descriptor_homeschool, school_descriptor_hybrid, school_descriptor_microschool, school_descriptor_other, school_descriptor_other_text) %>%
  mutate(homeschool = ifelse(school_descriptor_homeschool == 1, 1, 0),
         hybrid = ifelse(school_descriptor_hybrid == 1, 1, 0),
         microschool = ifelse(school_descriptor_microschool == 1, 1, 0),
         other = ifelse(school_descriptor_other == 1, 1, 0),
         other_text = ifelse(!is.na(school_descriptor_other_text), 1, 0)) %>%
  group_by(school_descriptor) %>%
  summarize(N = n(),
            homeschool = sum(homeschool),
            hybrid = sum(hybrid),
            microschool = sum(microschool),
            other = sum(other),
            other_text = sum(other_text)) %>%
  mutate(School_Type = recode(school_descriptor, "1" = "Charter", "2" = "District", "3" = "Independent")) %>%
  select(School_Type, N, homeschool, hybrid, microschool, other, other_text)

DT::datatable(type)

type_pivot <- type %>%
  pivot_longer(cols = c(homeschool, hybrid, microschool, other, other_text), 
               names_to = "descriptor", values_to = "count") %>%
  group_by(School_Type, descriptor) %>%
  summarize(total_count = sum(count))

DT::datatable(type_pivot)

```

