---
title: "Longitudinal Analysis"
author: "Janette Avelar"
date: '2023-03-22'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
library(tidyverse)
library(DT)
#data
load(here("data", "complete_canopy_2023.Rdata"))
#branding
# source(here("scripts", "branding.R"))
# theme_update(plot.title = element_text(hjust = 0.5))
dat <- full
```

#How many schools participated in both SY2021-22 and SY2022-23?

73 schools from last year (2021-22) participated this year (2022-23). This accounts for about 30% of all schools that responded this year. The full list of schools is displayed in the table below.

```{r schools since last year}
#pull in all datasets with school IDs and names first
merge_2019 <- import(here("data", "previous_datasets", "schools_2018-19.csv")) %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2018-19", 173))
merge_2021a <- import(here("data", "previous_datasets", "schools_2020-21a.csv")) %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2020-21", 144))
merge_2021b <- import(here("data", "previous_datasets", "schools_2020-21b.csv")) %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2020-21", 88))
merge_2021c <- import(here("data", "previous_datasets", "schools_2020-21c.csv")) %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2020-21", 99))
merge_2022 <- import(here("data", "previous_datasets", "schools_2021-22.csv")) %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2021-22", 161))
merge_2023 <- dat %>% 
  select(school_id, school_name) %>% 
  mutate(year = rep("2022-23", 251))
#merge last year and this year
repeat_schls_2023 <- merge_2023 %>% 
  left_join(merge_2022, by = c("school_id", "school_name")) %>% 
  filter(!is.na(year.y)) %>% 
  select(!c(year.x, year.y)) %>% 
  rename(`School ID` = school_id,
         `School Name` = school_name)
#table
datatable(repeat_schls_2023)
```

#How many schools participated in SY2020-21, SY2021-22, and SY2022-23?

There are 41 schools that have participated every year since AY2021-21. That's 7.5% of all schools we've had participate over time (N = 542). The list of schools is displayed in the table below.

```{r schools since COVID}
recent_schls <- merge_2021a %>% 
  #start by merging other 2021 datasets with a full merge to create full set
  full_join(merge_2021b, by = c("school_id", "school_name", "year")) %>% 
  full_join(merge_2021c, by = c("school_id", "school_name", "year")) %>% 
  #get rid of repeat schools
  unique() %>% 
  #left join following year to find crossover
  left_join(merge_2022, by = c("school_id", "school_name")) %>% 
  filter(!is.na(year.y)) %>% 
  left_join(merge_2023, by = c("school_id", "school_name")) %>% 
  filter(!is.na(year)) %>% 
  select(!starts_with("year")) %>% 
  rename(`School ID` = school_id,
         `School Name` = school_name)
#table
datatable(recent_schls)
```

#How many schools have participated across years?

We have a total of 542 Canopy schools, and their participation over time is displayed in the table below by year.

**Hint** If you want to sort by multiple rows, select 1 column followed by the next column you'd like to sort. If you select all 4 columns, it will display the 21 schools that have participated since the start of Canopy at the top of the table.

```{r}
long_schls <- merge_2019 %>% 
  full_join(merge_2021a, by = c("school_id", "school_name", "year")) %>% 
  full_join(merge_2021b, by = c("school_id", "school_name", "year")) %>% 
  full_join(merge_2021c, by = c("school_id", "school_name", "year")) %>% 
  full_join(merge_2022, by = c("school_id", "school_name", "year")) %>% 
  full_join(merge_2023, by = c("school_id", "school_name", "year")) %>% 
  mutate(wave = rep("Participated", length(817))) %>% 
  pivot_wider(names_from = year,
              values_from = wave) %>% 
  select(!school_id) %>% 
  rename(`School Name` = school_name)
datatable(long_schls)
```


#What practices are being adopted or retired over time? 

#How many tags have been retained across years?