---
title: "Modeling"
author: "Anwesha Guha"
date: "5/18/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library, include = FALSE}
library(pacman)
p_load(here, purrr, tidyverse, ggcorrplot, proxy, plotly, patchwork, psych, GPArotation, parameters, DT, nFactors)
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```


# Models

## Controlling for Race and SES  

The first model uses percentage of black students, hispanic students, locale, as well as free/reduced price lunch percent, students with disabilities percent, and percent of English learners.  

The plot details the top 10 absolute effects for each demographic variable listed above, as drawn from Gregor's work.

```{r data prep}
mod_prep <- full %>% 
  select(school_id, starts_with("practices_"), starts_with("self_reported")) %>% 
  select(-c(starts_with("self_reported_grades"), self_reported_locale_multiple, self_reported_ela, self_reported_math, starts_with("self_reported_locale_other"), ends_with("none_collected"))) %>% 
  pivot_longer(cols = c(starts_with("practices")),
               names_to = "tag",
               values_to = "value")

mod_prep <- mod_prep %>% 
  mutate(black_percent = self_reported_race_black/self_reported_total_enrollment, 
         hispanic_percent = self_reported_race_hispanic/self_reported_total_enrollment,
         frpl_percent = self_reported_frpl/self_reported_total_enrollment,
         swd_percent = self_reported_swd,
         ell_percent = self_reported_ell) %>% 
  rename(locale_urban = self_reported_locale_urban,
         locale_rural = self_reported_locale_rural,
         locale_suburban = self_reported_locale_suburban)

##from Gregor's code:
#first will run black percent and hispanic percent as controls
one_vars = c("self_reported_total_enrollment", "black_percent", "hispanic_percent", "locale_urban", "locale_rural", "locale_suburban", "frpl_percent", "swd_percent", "ell_percent")

logistic_one_dat = 
  mod_prep %>%
  select(value, tag, one_of(one_vars)) %>%
  na.omit() %>%
  group_by(tag) %>%
  # scaling non-binary predictors
  mutate_at(vars(
    c(
      "self_reported_total_enrollment",
      "black_percent",
      "hispanic_percent",
      "frpl_percent",
      "swd_percent",
      "ell_percent"
    )
  ), scale)

one_form = as.formula(paste("value", "~", paste(one_vars, collapse = "+")))
#plot(xm, "areas", prob = 0.9, prob_outer = 1)

bayes_mods = list()

library(scales)
library(rstanarm)
library(broom.mixed)
library(gmodels)

for(this_tag in unique(mod_prep$tag)) {
  bayes_mods[[this_tag]] = stan_glm(
    one_form,
    data = filter(logistic_one_dat, tag == this_tag),
    family = binomial(link = "logit"),
    prior = student_t( #df = n - 2
      df = 7,
      location = 0,
      scale = 2.5
    )
  )
  
}

bayes_tidy = lapply(bayes_mods, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") #%>%
  # mutate(
  #   nice_tag = label_tags(response),
  #   nice_demog = factor(label_dems(term))
  # )

## top 10 absolute effects per demo:

bayes_facet_dat = bayes_tidy %>% group_by(term) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10)

bayes_facet = 
  ggplot(bayes_facet_dat,
    aes(y = exp(estimate),
        x = reorder_within(response, estimate, within = term, fun = mean),
    )) +
  geom_col() + 
  labs(y = "Odds multiplier",
       x = "", 
       #fill = "Demographic",
       title = "Top 10 tags most/least related to each demographic") +
  facet_wrap(~ term, scales = "free_y") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .125, .25, .5, 1, 2, 4, 8, 16),
    labels = c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = 8), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))

bayes_facet
```

## Controlling for Aggregate BIPOC Population -- stopped here

```{r model 2 - bipoc, eval = FALSE}
#also run bipoc percent (all non-white)?
mod_prep2 <- clean_data_full %>% 
  select(school_id, starts_with("practices_"), locale, ends_with("_percent"), self_reported_total_enrollment) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "value") %>% 
  mutate(locale_urban = ifelse(locale=="Urban", 1, 0),
         locale_rural = ifelse(locale=="Rural", 1, 0),
         locale_suburban = ifelse(locale=="Suburban", 1, 0),
         non_white_percent = 1 - white_percent) %>% 
  select(-c(locale, aian_percent, asian_percent, black_percent, hispanic_percent, nhpi_percent, multiple_percent))

##from Gregor's code:
#first will run black percent and hispanic percent as controls
one_vars2 = c("self_reported_total_enrollment", "non_white_percent", "locale_urban", "locale_rural", "locale_suburban", "frpl_percent", "swd_percent", "ell_percent")

logistic_one_dat2 = 
  mod_prep2 %>%
  select(value, tag, one_of(one_vars2)) %>%
  na.omit() %>%
  group_by(tag) %>%
  # scaling non-binary predictors
  mutate_at(vars(
    c(
      "self_reported_total_enrollment",
      "non_white_percent",
      "frpl_percent",
      "swd_percent",
      "ell_percent"
    )
  ), scale)

one_form2 = as.formula(paste("value", "~", paste(one_vars2, collapse = "+")))
#plot(xm, "areas", prob = 0.9, prob_outer = 1)

bayes_mods2 = list()

for(this_tag in unique(mod_prep2$tag)) {
  bayes_mods2[[this_tag]] = stan_glm(
    one_form2,
    data = filter(logistic_one_dat2, tag == this_tag),
    family = binomial(link = "logit"),
    prior = student_t(
      df = 7,
      location = 0,
      scale = 2.5
    )
  )
  
}

bayes_tidy2 = lapply(bayes_mods2, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") %>%
  mutate(
    nice_tag = label_tags(response),
    nice_demog = factor(label_dems(term))
  )

## top 10 absolute effects per demo:

bayes_facet_dat2 = bayes_tidy2 %>% group_by(nice_demog) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10)

bayes_facet2 = 
  ggplot(bayes_facet_dat2,
    aes(y = exp(estimate),
        x = reorder_within(nice_tag, estimate, within = nice_demog, fun = mean),
    )) +
  geom_col() + 
  labs(y = "Odds multiplier",
       x = "", 
       #fill = "Demographic",
       title = "Top 10 tags most/least related to each demographic") +
  facet_wrap(~ nice_demog, scales = "free_y") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.125, .25, .5, 1, 2, 4, 8),
    labels = c("1/8", "1/4", "1/2", "1", "2", "4", "8"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = 8), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))

bayes_facet2
```

## Alternate Model

*also coming out weird, so ignore this*

```{r glm log regression locale, eval = FALSE}
########## ALL COMING OUT WEIRD, IGNORE ##################
#black and hispanic percent
glm_prep <- clean_data_full %>% 
  select(school_id, starts_with("practices_"), locale, ends_with("_percent"), self_reported_total_enrollment) %>% 
  mutate(non_white_percent = 1 - white_percent,
         locale = recode(locale,
           Urban = "_urban",
           Suburban = "_suburban",
           Rural = "_rural"
         )) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "value") %>% 
  #na.omit() %>% #kept in because it was used in previous modeling, but may reconsider
  group_by(tag) %>% 
  #scaling non-binary predictors
  mutate_at(vars(c(ends_with("_percent"))), scale)
######  model to build from ########
glm1 <- glm(value ~ 1 + black_percent + hispanic_percent + frpl_percent, 
            data = glm_prep, family = "binomial")
summary(glm1)
#vector for predictor vars
glm1_vars = c("black_percent", "hispanic_percent", "locale", "frpl_percent")
#use vector to build model for loop
glm1_form = as.formula(paste("value", "~", paste(glm1_vars, collapse = "+")))

glm1_mods = list()

for(this_tag in unique(glm_prep$tag)) {
  glm1_mods[[this_tag]] = glm(
    glm1_form,
    data = filter(glm_prep, tag == this_tag),
    family = "binomial")
  
}

glm1_tidy = lapply(glm1_mods, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") %>%
  mutate(
    nice_tag = label_tags(response),
    nice_demog = factor(label_dems(term))
  )

## top 10 absolute effects per demo:
glm1_facet_dat = glm1_tidy %>% group_by(nice_demog) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10)

glm1_facet = 
  ggplot(glm1_facet_dat,
    aes(y = exp(estimate),
        x = reorder_within(nice_tag, estimate, within = nice_demog, fun = mean),
    )) +
  geom_col() + 
  labs(y = "Odds multiplier",
       x = "", 
       #fill = "Demographic",
       title = "Top 10 tags most/least related to each demographic") +
  facet_wrap(~ nice_demog, scales = "free_y") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .125, .25, .5, 1, 2, 4, 8, 16),
    labels = c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = 8), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))

glm1_facet

#bipoc percent
```


```{r save model plots, eval = FALSE}
#black and hispanic percent as variable in odds model
ggsave("locale-odds-model-top10.png", plot = bayes_facet, path = here("output", "locale analysis"), 
       width = 20, height = 12, units = "in")
#nonwhite/bipoc percent as variable in odds model
ggsave("locale-bipoc-odds-model-top10.png", plot = bayes_facet2, path = here("output", "locale analysis"), 
       width = 20, height = 12, units = "in")
```
