---
title: "Model & Condition Innovation (67,68,71,73)"
author: "Merly Klaas"
date: "2023-06-01"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
pacman::p_load(rio, here, tidyverse, ggrepel, DT, forcats)
```

```{r}

#source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
model <- full %>% 
  select(school_id, 
         school_name,
         school_state,
         "Big Picture \nLearning"=model_usage_bpl,
         "EL Edu"=   model_usage_ele ,
         "IB" = model_usage_ib,
         "New Tech \nNetwork" = model_usage_ntn,
         "Opportunity \nCulture" =  model_usage_oc,
         "Summit \nLearning" = model_usage_sl,
         grades_high,
         self_reported_locale_rural,
         self_reported_locale_suburban,
         self_reported_locale_urban,
         self_reported_locale_multiple,
         waiver_application,
         political_impact)

cond <- full %>% 
  select(school_id,
         school_name,
         waiver_application)
```

```{r include=FALSE}
#Temporary manual branding used code here as Open Sans font keep making trouble

# main Transcend colors
transcend_cols = c(
  "#1A4C81",
  "#59C3B4",
  "#EF464B",
  "#ADE0EE"
)

transcend_cols2 = c(
  "#BC2582",
  "#FFA630",
  "#FFDE42",
  "#99C24D",
  "#218380",
  "#D3B7D7"
)

transcend_grays = c(
  "#4D4D4F",
  "#9D9FA2",
  "#D1D3D4"
)

  theme_1 <- function() {theme(
    plot.title = element_text(family = "Bebas Neue", color = "black"),
    #plot.subtitle = element_text(family = "Open Sans", size = rel(0.8)),
    plot.background = element_blank(),
    axis.text.x = element_text(colour = "black"),
     axis.text.y= element_blank(),
    axis.title.y=element_blank(),
    axis.ticks.y = element_blank(),
    panel.border = element_rect(colour = transcend_grays[1], fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )
  }
  
  
  theme_2 <- function() {
   theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black", hjust=0.5),
    #plot.subtitle = element_text(family = "Open Sans", size = rel(0.8)),
    plot.background = element_blank(),
    panel.border = element_rect(colour = transcend_grays[1], fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8),
   axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x =element_blank(),
  )
    }
  

  theme_3 <-function () {
    theme(
    plot.title.position = "plot",
   plot.title = element_text(family = "Bebas Neue", color = "black", hjust=0.5),
    #plot.subtitle = element_text(family = "Open Sans", size = rel(0.8)),
    plot.background = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.border = element_rect(colour = transcend_grays[1], fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8)
  ) 
  }
  
  scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
  
```
Check missing values 
```{r}

model %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  count(models, response) %>% 
  datatable()
model %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  count(response) %>% 
  datatable()

```

## Adopted Model
### "How many schools reported implementing adoptable models?"

For this graph, I recoded the response of 1 (somewhat) & 2 (to a large extent) to be* **YES**, and response of 0 and missing values to be **NO**

```{r}
#Check whether or not school implemented any adoptable models
model %>% 
  mutate(across(c(4:9),
                ~ case_when(. > 0 ~ "Yes",TRUE ~ "No"))) %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  ggplot(aes(x=models, 
             fill=response))+
  geom_bar(stat = "count", position = "dodge") + 
   scale_fill_manual(name="",
   values = transcend_cols) + 
  geom_text(stat = "count", aes(label = after_stat(count)), position = position_dodge(width = 1), vjust = -1, size = 3) + 
  labs(title = "How many schools reported implementing adoptable models?",
       x="") +
  theme_1()

```



```{r}
model %>% 
mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0"))) %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..])),
            position = position_fill(vjust = 0.5),
            stat = "count",
            size = 3,
            color = "white")+
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "To what degree is your school utilizing any of the adoptable models listed below?",
       ) +
  theme_1() + 
  theme(legend.position = "bottom") 
```


### Based on Locale: Adopted models

```{r fig.height=8, fig.width=8}
model %>%
mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0"))) %>% 
  pivot_longer(cols = starts_with("self_"),
               names_to = "category", 
               values_to = "count",
               names_prefix= "self_reported_locale_") %>% 
  filter(count == "1") %>% 
   ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
    facet_wrap(~category) + 
    geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="white", size =3)+
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "To what degree is your school utilizing any of the adoptable models listed below?",
       ) +
  theme_1() + 
  theme(legend.position = "bottom") 
  

```

### Adopted Models : High school vs Non high school

```{r fig.height=6, fig.width=8}
model %>%
mutate_if(is.numeric , replace_na, replace = 0) %>% 
  pivot_longer(cols = 4:9, 
               names_to = "models",
               values_to = "response") %>% 
  mutate(response = factor(response, levels = c("2", "1", "0")),
         grades_high = recode(grades_high, 
                              "1" = "High School",
                              "0" = "Non High School")) %>% 
  ggplot(aes(x=models, 
             fill=as.factor(response)))+
  geom_bar(stat = "count", position = "fill") + 
   geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="white", size =3)+
  scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("2", "1", "0"),
                     labels = c("To a large extent", "Somewhat", "Not at all/NA")) +
  labs(title = "To what degree is your school utilizing any of the adoptable models listed below?",
       ) +
  facet_wrap(~grades_high) + 
  theme_1() + 
  theme(legend.position = "bottom") 
  
  
  
```


## Conditions of Innovation

The majority of school/system in this sample did not consider applying for a waiver. 

```{r}
#overall <- 
full %>% 
  select(waiver_application) %>% 
  filter(waiver_application != "4") %>% 
  drop_na() %>% 
  ggplot(aes(x ="", fill = as.factor(waiver_application)))+ 
  geom_bar(position = "fill", width = 1, color = "white") + 
  coord_polar(theta = "y", start = 180) +
  geom_text(aes(label = scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..])),
            round= 1,
            position = position_fill(vjust = 0.5),
            stat = "count",
            size = 3)+
  scale_fill_manual(name="",
   values = transcend_cols2,
    breaks=c("3","1","2","0"),
    labels = c("Consider, but did not apply",
                "Applied & Received",
                "Applied, but did not receive it",
                "No"))+
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Has your school or system ever considered applying for a waiver from a local or state policy or regulation?") +
  theme_3()

#ggsave("finding-14.png", plot = overall, path = here("final_products", "draft-findings"),
     # width = 8, height = 6, units = "in")
  
  
```

### Conditions of Innovation: Policy Waiver Application Figure By State



```{r fig.height=8, fig.width=8}
full %>% 
  select(school_state, waiver_application) %>% 
  filter(waiver_application != "4") %>% 
  drop_na() %>% 
  ggplot(aes(x = fct_reorder(school_state, waiver_application), fill = as.factor(waiver_application)))+ 
  geom_bar(position = "fill", width = 1, color = "white") + 
#  coord_polar(theta = "y", start = 180) +
 geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), color="black", size = 3)+
  scale_fill_manual(name="",
   values = transcend_cols2,
    breaks=c("3","1","2","0"),
    labels = c("Consider, but did not apply",
                "Applied & Received",
                "Applied, but did not receive it",
                "No"))+
              # "Not sure",
              # "NA")) +
   scale_y_reordered() +
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Has your school or system ever considered applying for a waiver from a local or state policy or regulation?") +
  coord_flip() + 
  theme_3()
```

### Conditions of Innovation: Policy Waiver Application by Locale


```{r}
locale_dat <- 
  model %>%
  pivot_longer(cols = starts_with("self_"),
               names_to = "category", 
               values_to = "count",
               names_prefix= "self_reported_locale_") %>% 
  filter(count == "1",
         waiver_application != "4") %>% 
  mutate(category=str_to_title(category),
         waiver_application = factor(waiver_application, levels = c("1", "2", "3", "0"),
                                     labels = c("Received", 
                                                "Applied, but not received",
                                                "Considered, but not applied", 
                                                "Not considered"))) %>% 
  drop_na()

locale <- locale_dat %>% 
  ggplot(aes(x = category, fill = waiver_application))+ 
  geom_bar(position = "fill", width = 0.5, color = "white") + 
 # coord_polar(theta = "y", start = 180) +
  geom_text(aes(label = scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..])),
            round = 1,
            position = position_fill(vjust = 0.5),
            stat = "count",
            size = 3)+
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[3], transcend_cols[4])) +
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Has your school or system ever considered applying for a waiver from a local or state policy or regulation?") +
  coord_flip()+
  theme_3()

#ggsave("finding-14b.png", plot = locale, path = here("final_products", "draft-findings"),
      #width = 8, height = 6, units = "in")

```

Recreate for cover essay
```{r}
locale2 <- locale_dat %>% 
  ggplot(aes(x = category, fill = waiver_application))+ 
  geom_bar(position = position_fill(reverse = TRUE), width = 0.5, color = "white") + 
 # coord_polar(theta = "y", start = 180) +
  geom_text(aes(label = scales::percent(..count../tapply(..count.., ..x.. ,sum)[..x..], accuracy = 1)),
            position = position_fill(vjust = 0.5, reverse = TRUE),
            stat = "count",
            size = 3)+
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[3], transcend_cols[4])) +
  scale_y_continuous(expand = c(0,0))+
  labs(title = "Waiver Application from a Local of state policy or regulation",
       legend = "") +
  coord_flip()+
  theme(legend.direction = "horizontal",
        legend.title = element_blank()) +
  theme_3()
```

```{r}
ggsave_transcend(locale2, "cover_figure10", dir = here("final_products/essay-charts/cover_figures"), fig_width = 12, fig_height = 8)
```


### How did schools respond to the question about political debates? 

```{r}
full %>%
    select(political_impact) %>% 
    drop_na() %>% 
    count(political_impact) %>% 
    mutate(perc = round(proportions(n) * 100, 1),
           res = str_c(n ,"\n", "(", perc, ")%"),
           political_impact = factor(political_impact, levels = c("5", "4", "3", "2", "1")),
           category = "yes") %>% 
  ggplot(aes( x= category, y=n, fill = as.factor(political_impact)))+ 
  geom_bar(position = "stack", stat= "identity", color = "white", width = 0.2) + 
  geom_text(aes(label=res),
                position=position_stack(vjust=0.5),
            size=3)+
  labs(title = "To what extent do you agree with the following statement? \n“Political debates are detracting from our learning environment’s ability to stay focused on our mission.”") + 
scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values=c("#FFA630", "#FFDE42", "#D1D3D4", "#99C24D","#218380"),
   breaks=c("1", "2", "3", "4", "5"),
   labels=c("Strongly Disagree",
            "Disagree",
            "Neither agree nor disagree",
            "Agree",
            "Strongly Agree"),
   name="")+
  coord_flip()+
  theme_2()
```

```{r}
full %>%
    select(political_impact) %>% 
    count(political_impact) %>% 
    mutate(perc = round(proportions(n) * 100, 1),
           political_impact=recode(political_impact,
                                   "1"="Strongly Disagree",
                                   "2"="Disagree",
                                   "3"="Neither agree nor disagree",
                                   "4"="Agree",
                                   "5"="Strongly Agree")) %>% 
  datatable(caption = "To what extent do you agree with the following statement? \n“Political debates are detracting from our learning environment’s ability to stay focused on our mission (with NA)", options = list(dom = 't'))

```
### Political Debates by State
```{r fig.height=8, fig.width=8}
full %>%
    select(school_state,political_impact) %>% 
  drop_na() %>% 
    mutate(
           political_impact = factor(political_impact, levels = c("5", "4", "3", "2", "1"))) %>%
  ggplot(aes( x=fct_reorder(school_state, as.numeric(political_impact)), fill=as.factor(political_impact)))+ 
  geom_bar(stat= "count", 
           position = "fill")+
   geom_text(aes(label = ..count..), stat = "count", position = position_fill(vjust = 0.5), size =3)+
  labs(title = "By State: To what extent do you agree with the following statement? \n“Political debates are detracting from our learning environment’s ability to stay focused on our mission.”") +
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values=c("#FFA630", "#FFDE42", "#D1D3D4", "#99C24D","#218380"),
   breaks=c("1", "2", "3", "4", "5"),
   labels=c("Strongly Disagree",
            "Disagree",
            "Neither agree nor disagree",
            "Agree",
            "Strongly Agree"),
   name="")+
  coord_flip() + 
  theme_2()
```
### Political Debates by Locale
```{r}
poli_locale <- model %>%
  pivot_longer(cols = starts_with("self_"),
               names_to = "category", 
               values_to = "count",
               names_prefix= "self_reported_locale_") %>% 
  mutate(category = str_to_title(category)) %>% 
  filter(count == "1") %>% 
    drop_na() %>% 
  group_by(category) %>% 
    count(political_impact) %>% 
    mutate(perc = round(proportions(n) * 100),
           res = str_c(perc, "%"),
           political_impact = factor(political_impact, levels = c("5", "4", "3", "2", "1"))) %>% 
  ggplot(aes( x= as.factor(category), y=n, fill = as.factor(political_impact)))+ 
  geom_bar(position = "fill", stat= "identity", width = 0.5, color = "white", ) + 
  geom_text(aes(label=res),
                position=position_fill(vjust=0.5),
            size=4)+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values=c("#1B4C81", "#ADE0EE", "#D1D3D4", "#EF454A","#228380"),
   breaks=c("1", "2", "3", "4", "5"),
   labels=c("Strongly Disagree",
            "Disagree",
            "Neither agree nor disagree",
            "Agree",
            "Strongly Agree"),
   name="")+
  labs(title = "Political debates are detracting from our learning environment’s ability to stay focused on our mission.") +
  coord_flip() +
  theme(legend.direction = "horizontal") +
  theme_2()


ggsave_transcend(poli_locale, "cover_figure11", dir = here("final_products/essay-charts/cover_figures"), fig_width = 12, fig_height = 8)
```


### Political Debates High School - Non high school
```{r}
full %>%
    select(school_state, grades_high, political_impact) %>% 
    drop_na() %>% 
  group_by(grades_high) %>% 
    count(political_impact) %>% 
    mutate(perc = round(proportions(n) * 100, 1),
           res = str_c(n , "(", perc, ")%"),
           political_impact = factor(political_impact, levels = c("5", "4", "3", "2", "1")),
           grades_high = recode(grades_high, 
                                "1"= "High School",
                                "0"= "Non High School")) %>% 
  ggplot(aes( x= as.factor(grades_high), y=n, fill = as.factor(political_impact)))+ 
  geom_bar(position = "fill", stat= "identity", width = 0.2, color = "white", ) + 
  geom_text(aes(label=res),
                position=position_fill(vjust=0.5),
            size=2.5)+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values=c("#FFA630", "#FFDE42", "#D1D3D4", "#99C24D","#218380"),
   breaks=c("1", "2", "3", "4", "5"),
   labels=c("Strongly Disagree",
            "Disagree",
            "Neither agree nor disagree",
            "Agree",
            "Strongly Agree"),
   name="")+
  labs(title = "By High School VS Non-High School: To what extent do you agree with the following statement? \n“Political debates are detracting from our learning environment’s ability to stay focused on our mission.”") +
  coord_flip() +
  theme_2()
```