---
title: "Modeling Updated"
author: "Anwesha Guha"
date: "6/16/2023"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library, include = FALSE}
library(pacman)
p_load(here, purrr, tidyverse, ggcorrplot, proxy, plotly, patchwork, psych, GPArotation, parameters, DT, nFactors, broom)
library(scales)
library(rstanarm)
library(broom.mixed)
library(gmodels)
source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
```


# Models

The plot details the top 10 absolute effects for each demographic variable listed above, as drawn from Gregor's work.

```{r data prep}
mod_prep <- full %>% 
  select(school_id, school_descriptor, starts_with("practices_"), starts_with("self_reported"), starts_with("grades_")) %>% 
  select(-c(starts_with("self_reported_grades"), self_reported_locale_multiple, self_reported_ela, self_reported_math, starts_with("self_reported_locale_other"), ends_with("none_collected"))) %>% 
  pivot_longer(cols = c(starts_with("practices")),
               names_to = "tag",
               values_to = "value")

mod_prep <- mod_prep %>% 
  mutate(black_percent = self_reported_race_black/self_reported_total_enrollment, 
         hispanic_percent = self_reported_race_hispanic/self_reported_total_enrollment,
         white_percent = self_reported_race_white/self_reported_total_enrollment,
         native_percent = self_reported_race_aian/self_reported_total_enrollment,
         asian_percent = self_reported_race_asian/self_reported_total_enrollment,
         frpl_percent = self_reported_frpl/self_reported_total_enrollment,
         swd_percent = self_reported_swd,
         ell_percent = self_reported_ell,
         school_descriptor = case_when(school_descriptor == 1 ~ "Public district",
                                       school_descriptor == 2 ~ "Public charter",
                                       school_descriptor == 3 ~ "Independent/private")) %>% 
  rename(locale_urban = self_reported_locale_urban,
         locale_rural = self_reported_locale_rural,
         locale_suburban = self_reported_locale_suburban)


one_vars = c("self_reported_total_enrollment", "black_percent", "hispanic_percent", "white_percent", "native_percent", "asian_percent", "locale_urban", "locale_rural", "locale_suburban", "frpl_percent", "swd_percent", "ell_percent", "school_descriptor", "grades_prek", "grades_elementary", "grades_middle", "grades_high")

## below is filtering on tags that have been selected at least 25 times
mod_prep %>% 
  group_by(tag, value) %>% 
  summarise(val = n()) %>% 
  arrange(val) %>% 
  filter(val <= 25)

## we see that low select tags are the following
low_select_tags <- c("practices_other_leaders", "practices_heritage_language", "practices_enriched_virtual",
                     "practices_translanguaging", "practices_flipped_classroom")

## remove them for our analysis
mod_prep <- mod_prep %>% 
  filter(!(tag %in% low_select_tags))


logistic_one_dat = 
  mod_prep %>%
  select(value, tag, one_of(one_vars)) %>%
  na.omit() %>%
  group_by(tag) %>%
  # scaling non-binary predictors
  mutate_at(vars(
    c(
      "self_reported_total_enrollment",
      "black_percent",
      "hispanic_percent",
      "white_percent",
      "native_percent",
      "asian_percent",
      "frpl_percent",
      "swd_percent",
      "ell_percent"
    )
  ), scale)

one_form = as.formula(paste("value", "~", paste(one_vars, collapse = "+")))

#plot(xm, "areas", prob = 0.9, prob_outer = 1)

bayes_mods = list()


for(this_tag in unique(mod_prep$tag)) {
  bayes_mods[[this_tag]] = stan_glm(
    one_form,
    data = filter(logistic_one_dat, tag == this_tag),
    family = binomial(link = "logit"),
    prior = student_t( #df and scale updated to have less extreme values
      df = 10,
      location = 0,
      scale = 2.5
    )
  )
  
}

bayes_tidy = lapply(bayes_mods, tidy) %>%
  bind_rows(.id = "response") %>%
  filter(term != "(Intercept)") %>%
  mutate(
    nice_tag = label_tags()(response),
    nice_demog = factor(label_dems(term))
  )
```

```{r}
## top 10 absolute effects per demo:

bayes_facet_dat = bayes_tidy %>% 
  group_by(term) %>%
  arrange(desc(abs(estimate))) %>%
  slice(1:10)

bayes_facet = 
  ggplot(bayes_facet_dat,
    aes(y = exp(estimate),
        x = reorder_within(response, estimate, within = term, fun = mean)
    )) +
  geom_col(fill = transcend_cols[1]) + 
  labs(y = "Odds multiplier",
       x = "", 
       fill = "Demographic",
       title = "Top 10 tags most/least related to each demographic") +
  facet_wrap(~ term, scales = "free_y") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .125, .25, .5, 1, 2, 4, 8, 16),
    labels = c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = 8), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))

bayes_facet
```

```{r}
# ggsave("finding-9.png", plot = bayes_facet, path = here("final_products", "draft-findings"),
#        width = 22, height = 9, units = "in")
```


Plot above is clunky, so I am separating into different plots:
```{r}
list_plot <- bayes_facet_dat %>% 
  split(.$term) %>% 
  map(~ggplot(., aes(y = exp(estimate), 
             x = reorder_within(response, estimate, within = term, fun = mean)
             )) +
  geom_col(fill = transcend_cols[1]) + 
  labs(y = "Odds multiplier",
       x = "", 
       #fill = "Demographic",
       title = "Top 10 tags most/least related to each demographic") +
  scale_y_continuous(
    trans = "log",
    breaks = c(.0625, .125, .25, .5, 1, 2, 4, 8, 16),
    labels = c("1/16", "1/8", "1/4", "1/2", "1", "2", "4", "8", "16"),
    expand = expansion(0, 0)
  ) +
  scale_x_reordered() + 
  coord_flip() +
  guides(fill = "none") +
  theme(axis.text = element_text(size = 8), strip.text = element_text(size = rel(0.6)),
        panel.grid.major.y = element_blank(),
        axis.text.x = element_text(angle = -90, vjust = 0.5, hjust = 0))
  )
```

```{r}
file_names <- stringr::str_c(names(list_plot), ".png")
```

Run code below only to resave files
```{r}
pwalk(list(file_names, list_plot), ggsave, path = "output/modeling_low_tag_removed/")
```

```{r}
list_plot
```


# MODELING: Estimate model for each tag prevalence using (relatively) complete set of expanatory variables for general use in asnwering "How do tagging patterns differ across X?"

# STATE: How do tagging patterns differ across states?

# TYPE: How do tagging patterns differ among charter, district, and independent schools?

# LOCALE: How do tagging patterns differ among rural, urban, and suburban schools?

# LEVEL: How do tagging patterns differ among elementary, middle, and high schools?

# DEMOGRAPHICS: How do tagging patterns differ by student demographics (%white, %Black, %Hispanic/Latino, %Asian, %Native)?

^ these questions are answered!

