---
title: "Data Release Analysis"
author: "Janette Avelar"
date: '2023-03-22'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
library(tidyverse)
library(DT)
library(scales)
library(statebins)
# test <- import(here("data", "raw-ip.csv"))
load(here("data", "complete_canopy_2023.Rdata"))
source(here("scripts", "branding.R"))
# theme_update(plot.title = element_text(hjust = 0.5))
dat <- full
```

# Q1: Which states are represented or not?

We had 3 unknowns in the data, which are resolved in the table and figure below.

```{r states, fig.height=9, fig.width=9}
# ("N states in sample
# List of those not represented
# List (if any) U.S. territories represented")
#pull unique states
samp_states <- dat %>% 
  select(school_id, school_state) %>% 
  mutate(value = rep(1, 251)) %>% 
  select(!school_id) %>% 
  group_by(school_state) %>% 
  summarize(N = sum(value)) %>% 
  mutate(school_state = case_when(school_state == "" ~ "Unknown",
                                TRUE ~ as.character(school_state))) %>% 
  rename(`State` = school_state)
#table
datatable(samp_states)
#viz
plot_states <- samp_states %>% 
  filter(!is.na(State)) %>% 
ggplot(., aes(reorder(State, -N), N)) +
  geom_col(fill = transcend_cols[1]) +
  scale_y_continuous(limits=c(0, 25), expand = c(0,0)) +
  coord_flip() +
  labs(title = "Canopy Learning Environments by State",
       x = "",
       y = "") +
  geom_text(aes(label = signif(N)), 
            nudge_y = .5, 
            color = transcend_cols[3],
            fontface = "bold",
            size = 3.5,
            family = "sans")
plot_states
# ggsave("finding-12a.png", plot = plot_states, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

In total: 42 states were represented + D.C.
No U.S. territories are represented this round.

*States not represented:*  
* Alaska  
* Mississippi   
* Montana  
* Nebraska  
* Oklahoma  
* Oregon  
* West Virginia  
* Wyoming  

Alternate visualization for state

```{r state bins}
#abbreviations list
st_crosswalk <- tibble(state = state.name) %>%
   bind_cols(tibble(abb = state.abb)) %>% 
   bind_rows(tibble(state = "District of Columbia", abb = "DC")) %>% 
  rename(State = state)
#missing state list
na_state <- data.frame("State" = c("Alaska", "Mississippi", "Montana", "Nebraska", "Nevada", "Oklahoma", "Oregon", "West Virginia", "Wyoming"),
                          "N" = rep(NA, 9)) %>% 
  #merge missing state with data
  bind_rows(samp_states) %>% 
  mutate(color = case_when(
    is.na(N) ~ paste0(transcend_cols[3]),
    TRUE ~ paste0(transcend_cols[1])
  )) %>% #ignore color col for now
  left_join(st_crosswalk, by = "State") %>% 
  group_by(State) %>% 
  mutate(label = paste0({abb},"\nN=",{N}),
         label = case_when(
           is.na(N) ~ paste0({abb},"\nN=0"),
           TRUE ~ as.character(label)
         ))
#build plot
state_bins <-
statebins(state_data = na_state,
          font_size = 5,
          state_border_col = "black",
          state_border_size = 1,
          state_col = "State",
          value_col = "N",
          labels = "label",
          round = TRUE,
          light_label = "white",
          dark_label = "white") +
  theme_transcend +
  scale_fill_gradient(low = transcend_cols[4],
                       high = transcend_cols[1]) +
  labs(title = "2022-23 Canopy Learning Environments in the U.S.",
       subtitle = "Darker color indicates more schools are found in that state.\nGray indicates no schools are found in that state") +
  theme(legend.position = "none",
        axis.title = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major = element_blank())
ggsave("finding-12-V3.png", plot = state_bins, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")



sb = ggplot_build(state_bins)
sb$data[[2]]$label = na_state$label[match(sb$data[[2]]$label, na_state$abb)]
sb = ggplot_gtable(sb)
png(filename = here("final_products", "draft-findings", "finding-12-V4.png"),
    width = 12, height = 8, units = "in", res = 300)
plot(sb)
dev.off()
```

```{r population, include = FALSE}
## population for each state in 2020
## source https://www.census.gov/data/tables/time-series/demo/popest/2020s-state-total.html
## (used the April 1, 2020 base estimates)
state_pop = read_tsv(file = "Alabama	5024356
Alaska	733378
Arizona	7151507
Arkansas	3011555
California	39538245
Colorado	5773733
Connecticut	3605942
Delaware	989957
District of Columbia	689546
Florida	21538226
Georgia	10711937
Hawaii	1455273
Idaho	1839092
Illinois	12812545
Indiana	6785668
Iowa	3190372
Kansas	2937847
Kentucky	4505893
Louisiana	4657749
Maine	1362341
Maryland	6177213
Massachusetts	7029949
Michigan	10077325
Minnesota	5706504
Mississippi	2961288
Missouri	6154920
Montana	1084197
Nebraska	1961489
Nevada	3104624
New Hampshire	1377518
New Jersey	9289031
New Mexico	2117527
New York	20201230
North Carolina	10439414
North Dakota	779091
Ohio	11799374
Oklahoma	3959346
Oregon	4237291
Pennsylvania	13002689
Rhode Island	1097371
South Carolina	5118429
South Dakota	886677
Tennessee	6910786
Texas	29145428
Utah	3271614
Vermont	643085
Virginia	8631384
Washington	7705247
West Virginia	1793755
Wisconsin	5893725
Wyoming	576837", col_names = c("state", "population"))
```

```{r per_capita}
na_state = left_join(na_state, state_pop, by = c("State" = "state"))
na_state = na_state |>
  mutate(
    canopy_per_M = N / population * 1e6,
    canopy_per_M_label = paste(abb, round(coalesce(canopy_per_M, 0), 1), sep = "\n")
  )

state_bins_percap <- statebins(
    state_data = na_state,
    font_size = 5,
    state_border_col = "black",
    state_border_size = 1,
    state_col = "State",
    value_col = "canopy_per_M",
    # labels = "label",
    round = TRUE,
    # light_label = "white",
    # dark_label = "white"
  ) +
  theme_statebins() +
  scale_fill_distiller(type = "div", palette = "RdYlBu", na.value = "gray40") +
  # scale_fill_gradient2(
  #   low = transcend_cols[1],
  #   mid = transcend_cols2[2],
  #   high = transcend_cols[3],
  #   midpoint = 2,
  #   trans = sqrt_trans(),
  #   limits = c(0, 13.1)
  # ) +
  labs(title = "2022-23 Canopy Learning Environments in the U.S.",
       subtitle = "per million population")
state_bins_percap

sbpc = ggplot_build(state_bins_percap)
sbpc$data[[2]]$label = na_state$canopy_per_M_label[match(sbpc$data[[2]]$label, na_state$abb)]
sbpc = ggplot_gtable(sbpc)
png(filename = here("final_products", "draft-findings", "finding-12-V5.png"),
    width = 12, height = 8, units = "in", res = 300)
plot(sbpc)
dev.off()
```

# Q2: What's the breakdown of schools in the data by locale?


```{r locale}
locale <- dat %>% 
  select(school_name, exclusive_locale) %>% 
  mutate(rate = rep(1, 251)) %>% 
  group_by(exclusive_locale) %>% 
  summarize(N = sum(rate)) %>% 
  rename(`Geographic Locale` = exclusive_locale)
#table
datatable(locale)
#viz
locale_plot <- locale %>% 
  mutate(pct = N/251) %>% 
ggplot(., aes(reorder(`Geographic Locale`, N), pct, fill = `Geographic Locale`)) +
  geom_col() +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  scale_fill_manual(values = transcend_cols) +
  labs(title = "Canopy Learning Environments by Geographic Region",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
ggsave("finding-11-V2.png", plot = locale_plot, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")
```

## How do schools that identified as Other describe their locale?

The following schools identified their locale as `Other` (N = 7). The plot above includes these 7 schools, categorized based on their description below and per agreement made via Slack on 5.19.23.

```{r locale - other}
#others
locale_other <- dat %>% 
  select(school_name, self_reported_locale_other_text) %>% 
  filter(self_reported_locale_other_text != "") %>% 
  rename("Self-identified locale" = self_reported_locale_other_text,
         "School Name" = school_name)
#table
datatable(locale_other)
```

Note that the numbers are slightly wonky if you read the descriptions that schools chose to include--though Gem Prep Nampa and SEEQS identified themselves as `Other`, what they disclosed does not actually categorize them differently. Thus, suburban should be 1 higher (N = 32 because of Gem Prep) and urban should be 1 higher (N = 142 because of SEEQS).


# Q3: What's the breakdown of schools in the data by level?


```{r level}
#(N schools across levels (elementary, middle, high))
level <- dat %>% 
  select(starts_with("grades_")) %>%
  rename("Prekindergarten" = grades_prek,
         "Elementary" = grades_elementary,
         "Middle" = grades_middle,
         "High" = grades_high) %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "School level",
               values_to = "N")
#table
datatable(level, rownames = FALSE)
#factor viz data to reorder columns and generate percent
level <- level %>% 
  mutate(`School level` = factor(`School level`,
                levels = c("Prekindergarten", "Elementary", "Middle", "High"),
                labels = c("Prekindergarten", "Elementary", "Middle", "High")),
         pct = N/251)
#Modified V2 plot - added subtitle, converted to pct from count, added pct labels
level_plot <-
ggplot(level, aes(`School level`, pct, fill = `School level`)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Canopy Learning Environments by Level",
       subtitle = "Note: Some schools serve grades that cross levels, making our total percentage greater than 100%.\nTotal Canopy learning environments = 251",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
ggsave("finding-1-V3.png", plot = level_plot, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")
```


# Q4: What's the breakdown of schools in the data by school type?


```{r type}
#(N schools across school types (charter, district, independent))
# NOTE TO SELF TO MAKE A FUNCTION FROM THE ACTION BELOW
type <- dat %>% 
  select(school_descriptor) %>% 
  mutate(value = rep(1, 251),
         school_descriptor = case_when(
           school_descriptor == 1 ~ "Public district school",
           school_descriptor == 2 ~ "Public charter school",
           school_descriptor == 3 ~ "Independent (private) school"
         )) %>% 
  group_by(school_descriptor) %>% 
  summarize(N = sum(value)) %>% 
  rename("School Type" = school_descriptor)
#table
datatable(type, rownames = FALSE)
#viz
#create viz data - add pct
type <- type %>% 
  mutate(pct = N/251)

ggplot(type, aes(reorder(`School Type`, pct), pct, fill = `School Type`)) +
  geom_col() +
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[4])) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Canopy Learning Environments by Type",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01, 
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
# ggsave("finding-2-V2.png", plot = type_plot, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

```{r type pie chart}
 type_plot2 <- 

#create pie chart
ggplot(type, aes(x = "", y = N, fill = `School Type`)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(label = str_wrap(paste(label_percent(accuracy = 1)(pct), `School Type`, sep = "\n"), 10)),
            position = position_stack(vjust = .5),
            color = "white",
            fontface = "bold",
            size = 4.5,
            family = "sans") +
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[3])) +
  theme(plot.title = element_text(family = "Bebas Neue", 
                                  color = "black",
                                  size = 30,
                                  hjust = 0.5),
        legend.position = "none",
        text = element_text(size = 25,
                            family = "sans",
                            color = transcend_grays[1])) +
  labs(title = "Canopy Learning Environments by Type",
       x = "",
       y = "")
type_plot2
ggsave("finding-2-V3.png", plot = type_plot2, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")
```


# Q5: What's the racial/ethnic breakdown of Canopy school leadership teams?


```{r school leadership}
#(Sample demographics - school leadership team)
lead_team <- dat %>% 
  select(leadership_diversity) %>% 
  mutate(value = rep(1, 251),
         leadership_diversity = case_when(
           leadership_diversity == 0 ~ "Prefer not to say",
           leadership_diversity == 1 ~ "0 - 24% people of color",
           leadership_diversity == 2 ~ "25 - 49% people of color",
           leadership_diversity == 3 ~ "50 - 74% people of color",
           leadership_diversity == 4 ~ "75 - 100% people of color",
           leadership_diversity == 5 ~ "Not sure"
         )) %>% 
  group_by(leadership_diversity) %>% 
  summarize(N = sum(value)) %>% 
  rename("School Leadership Team Diversity" = leadership_diversity)
#table
datatable(lead_team, rownames = FALSE)
```


# Q6: What are the frequencies of all practices?


```{r practices, warning = FALSE}
#(N and % selected practices)
prac <- dat %>% 
  select(starts_with("practices_")) %>% 
  sapply(., as.numeric) %>% 
  as.data.frame() %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "Practice",
               values_to = "N") %>% 
  group_by(Practice) %>% 
  mutate(Pct = paste0(round(N/251*100, 2), "%"))
#table
datatable(prac)
```


## How many practices did schools typically select?

Average over the whole sample is ~34 tags.

```{r}
avg_prac <- dat %>% 
  select(school_id, starts_with("practices_")) %>% 
  group_by(school_id) %>% 
  mutate(sum_tag = rowSums(across(starts_with("practices_")))) %>% 
  ungroup() %>% 
  select(!starts_with("practices_")) %>% 
  summarize(`Average N Tags Selected` = mean(sum_tag))
#table
datatable(avg_prac, rownames = FALSE)
```


# Q7: What are the frequencies of core practices?


```{r core}
#(N and % core practices)
core_prac <- dat %>% 
  select(starts_with("core_")) %>% 
    summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "Core practice",
               values_to = "N") %>% 
  group_by(`Core practice`) %>% 
  mutate(Pct = paste0(round(N/251*100, 2), "%"))
#table
datatable(core_prac)
```


# Q8: How long have schools been implementing each core practice?

Note: This plot looks messy because it's too large right now. Once we have clusters finalized I can create condensed plots by cluster, or alternately organize them in whichever ways makes most sense.

```{r core time, fig.height=7, fig.width=10}
#(N and % implementation time)
time_prac <- dat %>% 
  select(starts_with("time_")) %>% 
    pivot_longer(everything(),
               names_to = "Practice",
               values_to = "N") %>% 
  mutate(`Less than a year` = case_when(N == 1 ~ 1, TRUE ~ 0),
         `1-2 years` = case_when(N == 2 ~ 1, TRUE ~ 0),
         `3-4 years` = case_when(N == 3 ~ 1, TRUE ~ 0),
         `5+ years` = case_when(N == 4 ~ 1, TRUE ~ 0),
         Practice = sub("time_", "practices_", Practice)) %>%
  select(!N) %>% 
  group_by(Practice) %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE)))
#table
datatable(time_prac)
#viz
#need long format?
plot_time_prac <- time_prac %>% 
  pivot_longer(cols = c(`Less than a year`, `1-2 years`, `3-4 years`, `5+ years`),
               names_to = "Time",
               values_to = "N") %>% 
  mutate(Time = factor(Time, levels = c(
    "Less than a year",
    "1-2 years",
    "3-4 years",
    "5+ years"
  )))
ggplot(plot_time_prac, aes(reorder(Practice, -N), N, fill = Time)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(limits=c(0, 85), expand = c(0,0)) +
  labs(title = "Practices by Time Implemented",
       x = "",
       y = "") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1))
###plan to facet by cluster once cluster analysis presented
```

## Top 10

```{r top 10 core practices with time}
top_prac <- 
  plot_time_prac %>% 
  group_by(Practice) %>% 
  mutate(sum = sum(N),
         pct = N/sum,
         tot_pct = round(sum/251*100, 0)) %>% 
  ungroup() %>% 
  group_by(Practice, Time) %>% 
  arrange(desc(sum)) %>% 
  ungroup() %>% 
  slice_head(., n = 40) %>% 
  ggplot(., aes(reorder(Practice, -sum), N, fill = Time)) +
  geom_col() +
  theme(panel.grid.major.x = element_blank()) +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(limits=c(0, 100), expand = c(0,0)) +
  labs(title = "Top 10 Practices by Time Implemented",
       x = "",
       y = "") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = c(.8,.75), 
        legend.box.background = element_rect(color = transcend_grays[1], linewidth = 2),
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)
        )
top_prac
ggsave("finding-8a-V2.png", plot = top_prac, path = here("final_products", "draft-findings"),
       width = 8, height = 6, units = "in")
```

Top 10 Frequency - No Time

```{r top 10 freq without time}
top_prac2 <- 
  plot_time_prac %>% 
  group_by(Practice) %>% 
  mutate(sum = sum(N),
         pct = round(100*(sum/251)), 1) %>% 
  ungroup() %>% 
  select(!c(Time, N)) %>% 
  group_by(Practice) %>% 
  arrange(desc(sum)) %>% 
  ungroup() %>% 
  slice_head(., n = 40) %>%
  unique()
pct_labs = paste0(top_prac2$pct, "%")
top_prac2 <- top_prac2 %>% 
ggplot(., aes(reorder(Practice, -sum), sum)) +
  geom_col(fill = transcend_cols[1]) +
  theme(panel.grid.major.x = element_blank()) +
  scale_y_continuous(limits=c(0, 100), expand = c(0,0)) +
  labs(title = "Frequency of Canopy's Top 10 Core Practices",
       x = "",
       y = "Number of learning environments") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "none",
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, size = rel(1)),
        axis.text.y = element_text(size = rel(1))) +
    geom_text(aes(label = pct_labs),
            nudge_y = .5,
            vjust = 0,
            color = transcend_cols[3],
            fontface = "bold",
            size = 5,
            family = "sans")
top_prac2
ggsave("finding-8a-V3.png", plot = top_prac2, path = here("final_products", "draft-findings"),
       width = 8, height = 6, units = "in")
```


Alternate view focusing on implementation time

```{r top 10 implementation time chart}
top_prac3 <- 
  plot_time_prac %>% 
  group_by(Practice) %>% 
  mutate(sum = sum(N),
         pct = N/sum) %>% 
  ungroup() %>% 
  group_by(Practice, Time) %>% 
  arrange(desc(sum)) %>% 
  ungroup() %>% 
  slice_head(., n = 40) %>% 
  #manual rank order tags by 5+ years implementation
  mutate(ord_time = case_when(
    Practice == "practices_career_prep" ~ 10,
    Practice == "practices_community_partnerships" ~ 9,
    Practice == "practices_pbl" ~ 8,
    Practice == "practices_all_courses_designed_for_inclusion" ~ 7,
    Practice == "practices_learning_paths" ~ 6,
    Practice == "practices_design_marginalized" ~ 5,
    Practice == "practices_sel_integrated" ~ 4,
    Practice == "practices_competency_education" ~ 3,
    Practice == "practices_culturally_responsive" ~ 2,
    Practice == "practices_restorative" ~ 1
  )) %>% 
  ggplot(., aes(reorder(Practice, ord_time), pct, fill = Time)) +
  geom_col() +
  bar_y_scale_percent +
  theme(panel.grid.major.y = element_blank()) +
  scale_fill_manual(values = transcend_cols) +
  coord_flip() +
  labs(title = "Top 10 Core Practices by Time Implemented",
       x = "",
       y = "") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        text = element_text(size = 10),
        axis.text.y = element_text(size = rel(1))) +
  guides(fill = guide_legend(reverse = TRUE))
top_prac3
ggsave("finding-8b-V3.png", plot = top_prac3, path = here("final_products", "draft-findings"),
       width = 8, height = 6, units = "in")
```

Top 10 frequency comparing core and non-core practices

```{r top 10 core + all}
core_names = full %>% 
  select(starts_with("core"))
prac_names = full %>% 
  select(starts_with("practices"))
comp_core <- full %>% 
  select(starts_with("practices"), starts_with("core")) %>% 
  pivot_longer(cols = everything(),
               names_to = "tag",
               values_to = "value") %>% 
  mutate(type = case_when(
    tag %in% names(core_names) ~ "Core Practice",
    tag %in% names(prac_names) ~ "Practice")) %>% 
  mutate(tag = str_replace_all(tag, "core", "practices")) %>% 
  group_by(tag, type) %>% 
  summarize(sum = sum(value),
            pct = sum/251) %>% 
  arrange(desc(pct)) %>% 
  #manual select top 10 core practice - can't figure out function
  filter(tag == "practices_pbl" | #repeat
         tag == "practices_competency_education" |
         tag == "practices_culturally_responsive" | #repeat
         tag == "practices_community_partnerships" | #repeat
         tag == "practices_sel_integrated"| #repeat 
         tag == "practices_all_courses_designed_for_inclusion" | #repeat
         tag == "practices_career_prep" |
         tag == "practices_restorative" | #repeat
         tag == "practices_design_marginalized" | #repeat
         tag == "practices_learning_paths"
#did not repeat = mental health, assessments for deeper learning, multiple opportunities to demonstrate mastery
           ) %>%
  ungroup() %>% 
  #ranking in order of biggest -> smallest gap
  mutate(rank = case_when(
    tag == "practices_pbl" ~ 7,
    tag == "practices_competency_education" ~ 10,
    tag == "practices_culturally_responsive" ~ 5,
    tag == "practices_community_partnerships" ~ 6,
    tag == "practices_sel_integrated" ~ 3,
    tag == "practices_all_courses_designed_for_inclusion" ~ 4,
    tag == "practices_career_prep" ~ 9,
    tag == "practices_restorative" ~ 2,
    tag == "practices_design_marginalized" ~ 1,
    tag == "practices_learning_paths" ~ 8
    )) %>% 
#PBL, culturally responsive practices, and integrated SEL were in the top 5 for both school-wide practices and core practices
  ggplot(., aes(reorder(tag, rank), pct, fill = type)) +
  geom_col(position = "identity") + #position = "dodge"
  bar_y_scale_percent +
  scale_x_discrete(labels = label_tags(wrap = 22)) +
  # coord_flip() +
  theme(panel.grid.major.x = element_blank()) +
  scale_fill_manual(values = transcend_cols) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "frequency of top 10 <span style = 'color: #1A4C81;'>core practices</span> compared to<br><span style = 'color: #59C3B4;'>all selected practices</span>",
       subtitle = stringr::str_wrap("When Canopy schools select the practices they use, they may select up to 5 practices as 'core practices' central to their learning environment.", 75),
       x = "",
       y = "") +
  theme(plot.title = element_markdown()) +
  geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)),
            size = 5,
            family = "sans",
            fontface = "bold",
            nudge_y = .025,
            vjust = .4) +
  guides(fill = guide_legend(title = "Type of Practice"))
comp_core
ggsave("finding-8c-V2.png", plot = comp_core, path = here("final_products", "draft-findings"),
       width = 8, height = 6, units = "in")
#I'm not sure if this is going to work, either from building it or interpretation
#the graph looks pretty misleading because of how stark the differences are
```


```{r include = FALSE}
#how many schools are "established" with 3 or more core practices implemented for 4+ years?
implementation <- full %>% 
  select(school_id, starts_with("time")) %>% 
  pivot_longer(cols = starts_with("time"),
               names_to = "tag",
               values_to = "time") %>% 
  mutate(est = case_when(
    time == 0 ~ 0, #not a core tag
    time == 1 ~ 0, #set to 1 for tags less than a year
    time == 2 ~ 0, #1-2 years
    time == 3 ~ 0, #3-4 years
    time == 4 ~ 1, #5+ years
    time == 5 ~ 0 #not sure
  )) %>% #tags are est if implemented 5+ years
  group_by(school_id) %>% 
  mutate(n_core = sum(est)) %>% #n_core = total tags that are est
  ungroup() %>% 
  select(school_id, n_core) %>% 
  unique() %>% #limit to just school and n_core
  mutate(rate = rep(1, nrow(.))) %>% #create rate var to count schools
  group_by(n_core) %>% 
  summarize(sum = sum(rate), #n schools with each # of est core prac
            pct = sum/251) #pct of schools with each # of est core prac
```


# Q9: Which core practices have been implemented most since AY20 
i.e., What did schools start implementing post-COVID?

The average number of schools selecting a given core 3-4 years ago (roughly around COVID) was 3.84. Therefore, I dropped any practices that less than 4 schools selected in order for us to look at the "above average" values. The table will allow you to filter by ascending or descending order if you want to see which ones were at the very top.

This left us with 27 practices with the following `Top 5`:
* Project-based learning (N = 20)  
* Culturally responsive practices (N = 17)  
* Restorative practices (N = 16)  
* Competency based education (N = 15)  
* Multi-tiered systems of support (N = 10)  

For reference, a total of 214 schools indicated these practices were core practices implemented 3-4 years (i.e., percentage is pulled from total schools that selected a tag and indicated implementing 3-4 years ago, rather than the total number of schools N = 251).

```{r core since covid}
covid_core <- time_prac %>% 
  select(!c(`Less than a year`, `5+ years`, `1-2 years`)) %>% 
  filter(`3-4 years` > 4) %>% 
  mutate(Pct = paste0(round(`3-4 years`/sum(`3-4 years`)*100, 2), "%"))
#table
datatable(covid_core)
```

#testing branding
```{r}
tag_bar <- tags %>% 
  select(school_id, starts_with("practices_")) %>% 
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "selected") %>% 
  group_by(tag) %>% 
  summarize(n = sum(selected)) %>% 
  head(n = 4)

ggplot(tag_bar, aes(x = tag, y = n, fill = tag)) +
  geom_col() +
  scale_x_discrete(labels = label_tags(wrap = 10)) +
  scale_fill_manual(values = transcend_cols)
```

#Misc. - Ignore

##Descriptors

```{r descriptors by type}
# desc <- full %>% 
#   select(school_name, school_descriptor, school_descriptor_homeschool, school_descriptor_hybrid, school_descriptor_microschool, school_descriptor_sws, school_descriptor_virtual) %>% 
#   mutate(school_type = case_when(
#     school_descriptor == 1 ~ "Public district school",
#            school_descriptor == 2 ~ "Public charter school",
#            school_descriptor == 3 ~ "Independent (private) school"),
#     school_description = case_when(
#       school_descriptor_homeschool == 1 ~ "Homeschool",
#       school_descriptor_hybrid == 1 ~ "Hybrid",
#       school_descriptor_microschool == 1 ~ "Microschool",
#       school_descriptor_sws == 1 ~ "School within school",
#       school_descriptor_virtual == 1 ~ "Virtual"
#     )) %>% 
#   select(-c(starts_with("school_descriptor"), school_name)) %>% 
#   mutate(rate = rep(1, 251)) %>% 
#   group_by(school_type, school_description) %>% 
#   summarize(n = sum(rate)) %>% 
#   ungroup() %>% 
#   mutate(pct = n/)
```

```{r}
# lang_schools <- dat %>% 
#   select(school_name, school_id, practices_dual_language, practices_translanguaging, practices_heritage_language, practices_assessments_bilingual, student_experience) %>% 
#   filter(practices_dual_language == 1 |
#          practices_translanguaging == 1 |
#            practices_heritage_language == 1 |
#            practices_assessments_bilingual == 1)
# export(lang_schools, "language.csv")
```


Quick analysis of schools reporting waiver overall and by locale

```{r}
quick_waiver <-
  full %>% 
  select(waiver_application) %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  group_by(waiver_application) %>% 
  summarize(sum = sum(rate),
            pct = sum/251) %>% 
  mutate(waiver_application = case_when(
    waiver_application == 0 ~ "No",
    waiver_application == 1 ~ "Yes + Received",
    waiver_application == 2 ~ "Yes + Did not receive",
    waiver_application == 3 ~ "Considered",
    waiver_application == 4 ~ "Not sure"
  ))
quick_waiver <-
  full %>% 
  select(waiver_application, exclusive_locale) %>% 
  mutate(rate = rep(1, nrow(.))) %>% 
  group_by(waiver_application, exclusive_locale) %>% 
  summarize(sum = sum(rate),
            pct = sum/251) %>% 
  mutate(waiver_application = case_when(
    waiver_application == 0 ~ "No",
    waiver_application == 1 ~ "Yes",
    waiver_application == 2 ~ "Yes",
    waiver_application == 3 ~ "Considered",
    waiver_application == 4 ~ "Not sure"
  ))

# "0 = No
# 1 = Yes, we have received a waiver
# 2 = Yes, we applied for a waiver, but did not receive it
# 3 = Yes, we considered applying for a waiver, did not ultimately choose to apply
# 4 = Not sure"
```

