---
title: "Data Release Analysis"
author: "Janette Avelar"
date: '2023-03-22'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
library(tidyverse)
library(DT)
library(scales)
# test <- import(here("data", "raw-ip.csv"))
load(here("data", "complete_canopy_2023.Rdata"))
source(here("scripts", "branding.R"))
theme_update(plot.title = element_text(hjust = 0.5))
dat <- full
```

# Q1: Which states are represented or not?

We had 3 unknowns in the data, which are resolved in the table and figure below.

```{r states, fig.height=9, fig.width=9}
# ("N states in sample
# List of those not represented
# List (if any) U.S. territories represented")
#pull unique states
samp_states <- dat %>% 
  select(school_id, school_state) %>% 
  mutate(value = rep(1, 251)) %>% 
  select(!school_id) %>% 
  group_by(school_state) %>% 
  summarize(N = sum(value)) %>% 
  mutate(school_state = case_when(school_state == "" ~ "Unknown",
                                TRUE ~ as.character(school_state))) %>% 
  rename(`State` = school_state)
#table
datatable(samp_states)
#viz
plot_states <- samp_states %>% 
  filter(!is.na(State)) %>% 
ggplot(., aes(reorder(State, -N), N)) +
  geom_col(fill = transcend_cols[1]) +
  scale_y_continuous(limits=c(0, 25), expand = c(0,0)) +
  coord_flip() +
  labs(title = "Canopy Learning Environments by State",
       x = "",
       y = "") +
  geom_text(aes(label = signif(N)), 
            nudge_y = .5, 
            color = transcend_cols[3],
            fontface = "bold",
            size = 3.5,
            family = "sans")
plot_states
# ggsave("finding-12a.png", plot = plot_states, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

In total: 42 states were represented + D.C.
No U.S. territories are represented this round.

*States not represented:*  
* Alaska  
* Mississippi   
* Montana  
* Nebraska  
* Oklahoma  
* Oregon  
* West Virginia  
* Wyoming  


# Q2: What's the breakdown of schools in the data by locale?


```{r locale}
#(N schools per geographic locale (urban, suburban, rural, multiple))
# locale <- dat %>% 
#   select(self_reported_locale_rural, self_reported_locale_suburban, self_reported_locale_urban, self_reported_locale_other) %>% 
#   rename("rural" = self_reported_locale_rural,
#          "suburban" = self_reported_locale_suburban,
#          "urban" = self_reported_locale_urban,
#          "other" = self_reported_locale_other) %>% 
#   mutate(Rural = if_else(rural == 1 & suburban == 0 & urban == 0 & other == 0, 1, 0),
#          Suburban = if_else(suburban == 1 & rural == 0 & urban == 0 & other == 0, 1, 0),
#          Urban = if_else(urban == 1 & rural == 0 & suburban == 0 & other == 0, 1, 0),
#          Other = if_else(other == 1, 1, 0),
#          Mixed = case_when(rural == 1 & suburban == 1 ~ 1,
#                            rural == 1 & urban == 1 ~ 1,
#                            urban == 1 & suburban == 1 ~ 1,
#                            urban == 1 & suburban == 1 & rural == 1 ~ 1,
#                          TRUE ~ 0)) %>% 
#   select(!c(rural, suburban, urban, other)) %>% 
#   summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
#     pivot_longer(everything(),
#                names_to = "Geographic Locale",
#                values_to = "N")
# #table
# datatable(locale, rownames = FALSE)

# code above was prior to 5.19.23 changes to mutually exclusive locale categories
locale <- dat %>% 
  select(school_name, exclusive_locale) %>% 
  mutate(rate = rep(1, 251)) %>% 
  group_by(exclusive_locale) %>% 
  summarize(N = sum(rate)) %>% 
  rename(`Geographic Locale` = exclusive_locale)
#table
datatable(locale)
#viz
#locale_plot <-
ggplot(locale, aes(reorder(`Geographic Locale`, N), N, fill = `Geographic Locale`)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(limits=c(0, 150), expand = c(0,0)) +
  labs(title = "Canopy Learning Environments by Geographic Region",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = signif(N)), 
            nudge_y = 3, 
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
# ggsave("finding-11.png", plot = locale_plot, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

## How do schools that identified as Other describe their locale?

The following schools identified their locale as `Other` (N = 7). The plot above includes these 7 schools, categorized based on their description below and per agreement made via Slack on 5.19.23.

```{r locale - other}
#others
locale_other <- dat %>% 
  select(school_name, self_reported_locale_other_text) %>% 
  filter(self_reported_locale_other_text != "") %>% 
  rename("Self-identified locale" = self_reported_locale_other_text,
         "School Name" = school_name)
#table
datatable(locale_other)
```

Note that the numbers are slightly wonky if you read the descriptions that schools chose to include--though Gem Prep Nampa and SEEQS identified themselves as `Other`, what they disclosed does not actually categorize them differently. Thus, suburban should be 1 higher (N = 32 because of Gem Prep) and urban should be 1 higher (N = 142 because of SEEQS).


# Q3: What's the breakdown of schools in the data by level?


```{r level}
#(N schools across levels (elementary, middle, high))
level <- dat %>% 
  select(starts_with("grades_")) %>%
  rename("Prekindergarten" = grades_prek,
         "Elementary" = grades_elementary,
         "Middle" = grades_middle,
         "High" = grades_high) %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "School level",
               values_to = "N")
#table
datatable(level, rownames = FALSE)
#factor viz data to reorder columns and generate percent
level <- level %>% 
  mutate(`School level` = factor(`School level`,
                levels = c("Prekindergarten", "Elementary", "Middle", "High"),
                labels = c("Prekindergarten", "Elementary", "Middle", "High")),
         pct = N/251)
#Modified V2 plot - added subtitle, converted to pct from count, added pct labels
level_plot <-
ggplot(level, aes(`School level`, pct, fill = `School level`)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Canopy Learning Environments by Level",
       subtitle = "Note: Some schools serve grades that cross levels, making our total percentage greater than 100%.",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01,
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
ggsave("finding-1-V2.png", plot = level_plot, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")
```


# Q4: What's the breakdown of schools in the data by school type?


```{r type}
#(N schools across school types (charter, district, independent))
# NOTE TO SELF TO MAKE A FUNCTION FROM THE ACTION BELOW
type <- dat %>% 
  select(school_descriptor) %>% 
  mutate(value = rep(1, 251),
         school_descriptor = case_when(
           school_descriptor == 1 ~ "Public district school",
           school_descriptor == 2 ~ "Public charter school",
           school_descriptor == 3 ~ "Independent (private) school"
         )) %>% 
  group_by(school_descriptor) %>% 
  summarize(N = sum(value)) %>% 
  rename("School Type" = school_descriptor)
#table
datatable(type, rownames = FALSE)
#viz
#create viz data - add pct
type <- type %>% 
  mutate(pct = N/251)

ggplot(type, aes(reorder(`School Type`, pct), pct, fill = `School Type`)) +
  geom_col() +
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[4])) +
  bar_y_scale_percent +
  theme(panel.grid.major.x = element_blank()) +
  labs(title = "Canopy Learning Environments by Type",
       x = "",
       y = "") +
  theme(legend.position = "none") +
    geom_text(aes(label = scales::label_percent(accuracy = 1)(pct)), 
            nudge_y = .01, 
            vjust = 0,
            color = "black",
            fontface = "bold",
            size = 5,
            family = "sans")
# ggsave("finding-2-V2.png", plot = type_plot, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

```{r type pie chart}
 type_plot2 <- 

#create pie chart
ggplot(type, aes(x = "", y = N, fill = `School Type`)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(label = str_wrap(paste(label_percent(accuracy = 1)(pct), `School Type`, sep = "\n"), 10)),
            position = position_stack(vjust = .5),
            color = "white",
            fontface = "bold",
            size = 4.5,
            family = "sans") +
  scale_fill_manual(values = c(transcend_cols[1], transcend_cols[2], transcend_cols[3])) +
  theme(plot.title = element_text(family = "Bebas Neue", 
                                  color = "black",
                                  size = 30,
                                  hjust = 0.5),
        legend.position = "none",
        text = element_text(size = 14,
                            family = "sans",
                            color = transcend_grays[1])) +
  labs(title = "Canopy Learning Environments by Type",
       x = "",
       y = "")
type_plot2
ggsave("finding-2-V2.png", plot = type_plot2, path = here("final_products", "draft-findings"),
       width = 12, height = 8, units = "in")
```


# Q5: What's the racial/ethnic breakdown of Canopy school leadership teams?


```{r school leadership}
#(Sample demographics - school leadership team)
lead_team <- dat %>% 
  select(leadership_diversity) %>% 
  mutate(value = rep(1, 251),
         leadership_diversity = case_when(
           leadership_diversity == 0 ~ "Prefer not to say",
           leadership_diversity == 1 ~ "0 - 24% people of color",
           leadership_diversity == 2 ~ "25 - 49% people of color",
           leadership_diversity == 3 ~ "50 - 74% people of color",
           leadership_diversity == 4 ~ "75 - 100% people of color",
           leadership_diversity == 5 ~ "Not sure"
         )) %>% 
  group_by(leadership_diversity) %>% 
  summarize(N = sum(value)) %>% 
  rename("School Leadership Team Diversity" = leadership_diversity)
#table
datatable(lead_team, rownames = FALSE)
```


# Q6: What are the frequencies of all practices?


```{r practices, warning = FALSE}
#(N and % selected practices)
prac <- dat %>% 
  select(starts_with("practices_")) %>% 
  sapply(., as.numeric) %>% 
  as.data.frame() %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "Practice",
               values_to = "N") %>% 
  group_by(Practice) %>% 
  mutate(Pct = paste0(round(N/251*100, 2), "%"))
#table
datatable(prac)
```


## How many practices did schools typically select?

Average over the whole sample is ~34 tags.

```{r}
avg_prac <- dat %>% 
  select(school_id, starts_with("practices_")) %>% 
  group_by(school_id) %>% 
  mutate(sum_tag = rowSums(across(starts_with("practices_")))) %>% 
  ungroup() %>% 
  select(!starts_with("practices_")) %>% 
  summarize(`Average N Tags Selected` = mean(sum_tag))
#table
datatable(avg_prac, rownames = FALSE)
```


# Q7: What are the frequencies of core practices?


```{r core}
#(N and % core practices)
core_prac <- dat %>% 
  select(starts_with("core_")) %>% 
    summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(everything(),
               names_to = "Core practice",
               values_to = "N") %>% 
  group_by(`Core practice`) %>% 
  mutate(Pct = paste0(round(N/251*100, 2), "%"))
#table
datatable(core_prac)
```


# Q8: How long have schools been implementing each core practice?

Note: This plot looks messy because it's too large right now. Once we have clusters finalized I can create condensed plots by cluster, or alternately organize them in whichever ways makes most sense.

```{r core time, fig.height=7, fig.width=10}
#(N and % implementation time)
time_prac <- dat %>% 
  select(starts_with("time_")) %>% 
    pivot_longer(everything(),
               names_to = "Practice",
               values_to = "N") %>% 
  mutate(`Less than a year` = case_when(N == 1 ~ 1, TRUE ~ 0),
         `1-2 years` = case_when(N == 2 ~ 1, TRUE ~ 0),
         `3-4 years` = case_when(N == 3 ~ 1, TRUE ~ 0),
         `5+ years` = case_when(N == 4 ~ 1, TRUE ~ 0),
         Practice = sub("time_", "practices_", Practice)) %>%
  select(!N) %>% 
  group_by(Practice) %>% 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE)))
#table
datatable(time_prac)
#viz
#need long format?
plot_time_prac <- time_prac %>% 
  pivot_longer(cols = c(`Less than a year`, `1-2 years`, `3-4 years`, `5+ years`),
               names_to = "Time",
               values_to = "N") %>% 
  mutate(Time = factor(Time, levels = c(
    "Less than a year",
    "1-2 years",
    "3-4 years",
    "5+ years"
  )))
ggplot(plot_time_prac, aes(reorder(Practice, -N), N, fill = Time)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(limits=c(0, 85), expand = c(0,0)) +
  labs(title = "Practices by Time Implemented",
       x = "",
       y = "") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1))
###plan to facet by cluster once cluster analysis presented
```

## Top 10

```{r top 10 core practices with time}
top_prac <- 
  plot_time_prac %>% 
  group_by(Practice) %>% 
  mutate(sum = sum(N)) %>% 
  ungroup() %>% 
  group_by(Practice, Time) %>% 
  arrange(desc(sum)) %>% 
  ungroup() %>% 
  slice_head(., n = 40) %>% 
  ggplot(., aes(reorder(Practice, -sum), N, fill = Time)) +
  geom_col() +
  scale_fill_manual(values = transcend_cols) +
  scale_y_continuous(limits=c(0, 100), expand = c(0,0)) +
  labs(title = "Top 10 Practices by Time Implemented",
       x = "",
       y = "") +
  scale_x_discrete(labels = label_tags()) +
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        text = element_text(size = 7),
        axis.text.x = element_text(angle = 45, hjust = 1)) #+
    # geom_text(aes(y = Practice, label = signif(sum)),
    #         #nudge_y = .5,
    #         color = transcend_cols[3],
    #         fontface = "bold",
    #         size = 5,
    #         family = "sans")
top_prac
# ggsave("finding-8.png", plot = top_prac, path = here("final_products", "draft-findings"),
#        width = 8, height = 6, units = "in")
```


# Q9: Which core practices have been implemented most since AY20 
i.e., What did schools start implementing post-COVID?

The average number of schools selecting a given core 3-4 years ago (roughly around COVID) was 3.84. Therefore, I dropped any practices that less than 4 schools selected in order for us to look at the "above average" values. The table will allow you to filter by ascending or descending order if you want to see which ones were at the very top.

This left us with 27 practices with the following `Top 5`:
* Project-based learning (N = 20)  
* Culturally responsive practices (N = 17)  
* Restorative practices (N = 16)  
* Competency based education (N = 15)  
* Multi-tiered systems of support (N = 10)  

For reference, a total of 214 schools indicated these practices were core practices implemented 3-4 years (i.e., percentage is pulled from total schools that selected a tag and indicated implementing 3-4 years ago, rather than the total number of schools N = 251).

```{r core since covid}
covid_core <- time_prac %>% 
  select(!c(`Less than a year`, `5+ years`, `1-2 years`)) %>% 
  filter(`3-4 years` > 4) %>% 
  mutate(Pct = paste0(round(`3-4 years`/sum(`3-4 years`)*100, 2), "%"))
#table
datatable(covid_core)
```

#testing branding
```{r}
tag_bar <- tags %>% 
  select(school_id, starts_with("practices_")) %>% 
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "selected") %>% 
  group_by(tag) %>% 
  summarize(n = sum(selected)) %>% 
  head(n = 4)

ggplot(tag_bar, aes(x = tag, y = n, fill = tag)) +
  geom_col() +
  scale_x_discrete(labels = label_tags(wrap = 10)) +
  scale_fill_manual(values = transcend_cols)
```

#Misc. - Ignore

##Descriptors

```{r descriptors by type}
# desc <- full %>% 
#   select(school_name, school_descriptor, school_descriptor_homeschool, school_descriptor_hybrid, school_descriptor_microschool, school_descriptor_sws, school_descriptor_virtual) %>% 
#   mutate(school_type = case_when(
#     school_descriptor == 1 ~ "Public district school",
#            school_descriptor == 2 ~ "Public charter school",
#            school_descriptor == 3 ~ "Independent (private) school"),
#     school_description = case_when(
#       school_descriptor_homeschool == 1 ~ "Homeschool",
#       school_descriptor_hybrid == 1 ~ "Hybrid",
#       school_descriptor_microschool == 1 ~ "Microschool",
#       school_descriptor_sws == 1 ~ "School within school",
#       school_descriptor_virtual == 1 ~ "Virtual"
#     )) %>% 
#   select(-c(starts_with("school_descriptor"), school_name)) %>% 
#   mutate(rate = rep(1, 251)) %>% 
#   group_by(school_type, school_description) %>% 
#   summarize(n = sum(rate)) %>% 
#   ungroup() %>% 
#   mutate(pct = n/)
```

```{r}
# lang_schools <- dat %>% 
#   select(school_name, school_id, practices_dual_language, practices_translanguaging, practices_heritage_language, practices_assessments_bilingual, student_experience) %>% 
#   filter(practices_dual_language == 1 |
#          practices_translanguaging == 1 |
#            practices_heritage_language == 1 |
#            practices_assessments_bilingual == 1)
# export(lang_schools, "language.csv")
```

