---
html_document:
  toc: yes
  toc_float: yes
  theme: flatly
  highlight: tango
  code_folding: hide
author: "Karen Zyskind"
date: "2023-05-19"
title: "Differences in Types"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
load("~/Desktop/canopy-dac-23/data/complete_canopy_2023.RData")
library(tidyverse)
library(here)
library(reshape2)
library(cowplot)
library(dplyr)
library("pacman")
library("DT")
pacman::p_load("summarytools")
load(here("data", "complete_canopy_2023.RData"))

```

#Guiding questions/tasks & format:  
`SCHOOL TYPE`: How do tagging patterns differ among district, charter, independent, sws, virtual, microschool, homeschool, and hub schools?  
* table of biggest differences between categories  
* odds ratio  
* visualization 

##Type and how many tags
First recall Type and Pattern of Tags 
First looking at type count 
Charter n= 113
District n= 102
Independent = 36

```{r Type, echo=FALSE}
library(gt)

# Count the number of schools for each school descriptor
school_counts <- schools %>%
  count(school_descriptor)

# Load the necessary packages
library(gt)

# Create the table
school_counts_table <- school_counts %>%
  mutate(school_descriptor = c("Charter", "District", "Independent")) %>%
  gt() %>%
  tab_header(
    title = "Number of Schools by Descriptor"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = everything())
  )

# Display the table
school_counts_table




```

## Countsand differences table 

The values in the table indicate the count or frequency of each tagging practice for each school descriptor. For example, in the "Independent (private) school" row, the "practices_a_la_carte" column has a value of 5, indicating that this practice is present in 5 schools categorized as independent private schools. 

```{r Differences, echo=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)

# Merge the tags and schools data frames
merged_data <- inner_join(tags, schools, by = "school_id")

# Define the custom labels for school_descriptor
school_descriptor_labels <- c("Public district school", "Public charter school", "Independent (private) school")

# Filter necessary columns for the analysis and apply custom labels
filtered_data <- merged_data %>%
  select(starts_with("practices_"), school_descriptor) %>%
  mutate(school_descriptor = case_when(
    school_descriptor == 1 ~ school_descriptor_labels[1],
    school_descriptor == 2 ~ school_descriptor_labels[2],
    school_descriptor == 3 ~ school_descriptor_labels[3],
    TRUE ~ as.character(school_descriptor)  # Handle other values if present
  ))

# Create the tag_differences object
tag_differences <- filtered_data %>%
  group_by(school_descriptor) %>%
  summarise_all(sum)

# Calculate the absolute differences between counts for each tag
type_tbl <- tag_differences %>%
  mutate(dis_char_diff = abs(practices_a_la_carte - practices_accommodations),
         char_ind_diff = abs(practices_accommodations - practices_adult_wellness),
         ind_dis_diff = abs(practices_adult_wellness - practices_a_la_carte))

# Create the table
tag_differences_table <- type_tbl %>%
  kable(caption = "Differences in Tags by School Descriptor", align = "c") %>%
  kable_styling("striped")

# Display the table of differences
print(tag_differences_table)

```

## Differences in charter and independent schools 
practices_career_prep: Charter schools place a higher emphasis on career preparation programs compared to independent schools, with a difference of 35.

practices_makerspace: Charter schools prioritize the establishment of makerspaces for hands-on learning experiences more than independent schools, showing a difference of 32.
```{r charter and indep, echo=FALSE}

str(type_tbl)
# Calculate the differences in charter and independent schools
diff_career_prep <- type_tbl$practices_career_prep[2] - type_tbl$practices_career_prep[3]
diff_makerspace <- type_tbl$practices_makerspace[2] - type_tbl$practices_makerspace[3]

# Print the differences
cat("practices_career_prep: Charter schools place a higher emphasis on career preparation programs compared to independent schools, with a difference of", diff_career_prep, "\n")
cat("practices_makerspace: Charter schools prioritize the establishment of makerspaces for hands-on learning experiences more than independent schools, showing a difference of", diff_makerspace, "\n")

# Reshape the data to create a single "tag" column
char_ind_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with largest differences
char_ind_diff <- char_ind_diff %>%
  group_by(tag) %>%
  summarize(`Charter vs. Independent Difference` = abs(sum(value[2]) - sum(value[3]))) %>%
  top_n(n = 10, wt = `Charter vs. Independent Difference`) %>%
  arrange(desc(`Charter vs. Independent Difference`))

# Create the bar chart visualization
char_ind_viz <- char_ind_diff %>%
  ggplot(aes(reorder(tag, `Charter vs. Independent Difference`), `Charter vs. Independent Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference", 
       title = "Top 10 Tags with Largest Differences between Charter and Independent Schools") +
  coord_flip()

# Display the visualization
char_ind_viz

# Print the table
print(char_ind_diff)



```

## Large Differences in Independent Vs District 
practices_mtss_academics: Independent schools exhibit a higher presence on Multi-Tiered System of Supports (MTSS) for academics compared to district schools, with a difference of 69.
practices_career_prep: Independent schools prioritize career preparation programs more than district schools, showing a difference of 66.

```{r Indep Dist, echo=FALSE}
# Reshape the data to create a single "tag" column
ind_dis_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with largest differences
ind_dis_diff <- ind_dis_diff %>%
  group_by(tag) %>%
  summarize(`Independent vs. District Difference` = abs(sum(value[1]) - sum(value[3]))) %>%
  top_n(n = 10, wt = `Independent vs. District Difference`) %>%
  arrange(desc(`Independent vs. District Difference`))

# Create the bar chart visualization
ind_dis_viz <- ind_dis_diff %>%
  ggplot(aes(reorder(tag, `Independent vs. District Difference`), `Independent vs. District Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference", 
       title = "Top 10 Tags with Largest Differences between Independent and District Schools") +
  coord_flip()

# Display the visualization
ind_dis_viz

# Print the table
print(ind_dis_diff)


```
## Largest difference charter vs District 
# Table Differences largest Differences 
This looked at the 10 tags with the largest differences between District and Charter schools. Findings showed variations in practices across various areas. For example, practices related to MTSS academics showed a difference of 63 points between District and Charter schools. Similarly, culturally responsive practices had a difference of 59 points, assessments had a difference of 57 points, and so on. 
```{r, chartervdistric, echo=FALSE}

# Reshape the data to create a single "tag" column
char_dis_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with largest differences
char_dis_diff <- char_dis_diff %>%
  group_by(tag) %>%
  summarize(`Charter vs. District Difference` = abs(sum(value[2]) - sum(value[3]))) %>%
  top_n(n = 10, wt = `Charter vs. District Difference`) %>%
  arrange(desc(`Charter vs. District Difference`))

# Create the bar chart visualization
char_dis_viz <- char_dis_diff %>%
  ggplot(aes(reorder(tag, `Charter vs. District Difference`), `Charter vs. District Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference", 
       title = "Top 10 Tags with Largest Differences between Charter and District Schools") +
  coord_flip()

# Display the visualization
char_dis_viz

# Print the table
print(char_dis_diff)



```


## Smallest differences independent and district 
```{r small indept, echo=FALSE}


# Reshape the data to create a single "tag" column
ind_dis_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with smallest differences
ind_dis_diff <- ind_dis_diff %>%
  group_by(tag) %>%
  summarize(`Independent vs. District Difference` = abs(sum(value[1]) - sum(value[3]))) %>%
  top_n(n = 10, wt = `Independent vs. District Difference`) %>%
  arrange(`Independent vs. District Difference`)

# Create the bar chart visualization
ind_dis_viz <- ind_dis_diff %>%
  ggplot(aes(reorder(tag, `Independent vs. District Difference`), `Independent vs. District Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference",
       title = "Top 10 Tags with Smallest Differences between Independent and District Schools") +
  coord_flip()

# Display the visualization
ind_dis_viz

# Print the table
print(ind_dis_diff)


```

## Smallest difference district and charter
```{r small district, echo=FALSE}
# Reshape the data to create a single "tag" column
dis_char_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with smallest differences
dis_char_diff <- dis_char_diff %>%
  group_by(tag) %>%
  summarize(`District vs. Charter Difference` = abs(sum(value[1]) - sum(value[2]))) %>%
  top_n(n = 10, wt = `District vs. Charter Difference`) %>%
  arrange(`District vs. Charter Difference`)

# Create the bar chart visualization
dis_char_viz <- dis_char_diff %>%
  ggplot(aes(reorder(tag, `District vs. Charter Difference`), `District vs. Charter Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference",
       title = "Top 10 Tags with Smallest Differences between District and Charter Schools") +
  coord_flip()

# Display the visualization
dis_char_viz

# Print the table
print(dis_char_diff)

```

## Smallest difference charter vs independent 
```{r small chart and ind, echo=FALSE}
# Reshape the data to create a single "tag" column
char_ind_diff <- type_tbl %>%
  pivot_longer(cols = starts_with("practices_"),
               names_to = "tag",
               values_to = "value") %>%
  filter(!is.na(value))

# Select top 10 tags with smallest differences
char_ind_diff <- char_ind_diff %>%
  group_by(tag) %>%
  summarize(`Charter vs. Independent Difference` = abs(sum(value[2]) - sum(value[1]))) %>%
  top_n(n = 10, wt = `Charter vs. Independent Difference`) %>%
  arrange(`Charter vs. Independent Difference`)

# Create the bar chart visualization
char_ind_viz <- char_ind_diff %>%
  ggplot(aes(reorder(tag, `Charter vs. Independent Difference`), `Charter vs. Independent Difference`)) +
  geom_bar(stat = "identity", fill = "dodgerblue") +
  theme_minimal() +
  labs(x = "Tag", y = "Difference",
       title = "Top 10 Tags with Smallest Differences between Charter and Independent Schools") +
  coord_flip()

# Display the visualization
char_ind_viz

# Print the table
print(char_ind_diff)



```
