---
title: "School Leader Analysis"
author: "Anwesha Guha"
date: '2023-04-23'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(here, rio, tidyverse, DT, janitor)

load(here("data", "complete_canopy_2023.RData"))
```

Where are schools innovating?

# Setup

Let's see what variables we are working with.
```{r}
confidential_names <- tibble(variable.names(confidential))
```


# Q1: What are the racial/ethnic and gender identities of the primary school leader? (How many schools are led by school leaders of different racial/ethnic identities?)


## Gender

```{r}
# Renaming for visualization/interpretation
leader_char <- confidential %>% 
  mutate(leader1_gender = recode(leader1_gender,
            "1" = "Man",
            "2" = "Woman",
            "5" = as.character(leader1_gender_self_identify_text),
            "0" = "Prefer not to say",
            "3" = "Non-binary/gender-nonconforming",
            "1,2" = "Man and/or Woman",
            "1,2,3,4" = "All and/or Transgender"
  ))

#looks like self-identify text might have been reported in the wrong column, so recoding that one manually
leader_char$leader1_gender[leader_char$leader1_gender_self_identify_text == 1] <- "Man"
```

```{r}
leader_char %>% 
  group_by(leader1_gender) %>% 
  summarise(n = n()) %>% 
  datatable()
```

```{r}
gender_plot <- leader_char %>% 
  group_by(leader1_gender) %>% 
  summarise(n = n()) %>% 
  ggplot(aes(leader1_gender, n)) +
  geom_bar(stat = "identity", aes(fill = leader1_gender), show.legend = FALSE) +
  theme_minimal() +
  theme(axis.text.x = element_blank()) +
  geom_text(aes(label=n), vjust=0) +
  geom_text(aes(label=leader1_gender), vjust=-1.5)  +
  scale_fill_viridis_d()

gender_plot
```

## Race

```{r}
leader_char <- leader_char %>% 
  mutate(leader1_race = recode(leader1_race,
            "0" = "Prefer not to say",                               
            "1" = "American Indian or Alaska Native",
            "2" = "Asian",
            "3" = "Black or African American",
            "4" = "Hispanic or Latinx",
            "5" = "Native Hawaiian or Pacific Islander",
            "6" = "White",
            "7" = as.character(leader1_race_self_identify_text),
            "1,2,4,6" = "American Indian or Alaska Native, Asian, Hispanic or Latinx, White",
            "1,3" = "American Indian or Alaska Native, Black or African American",
            "1,3,4,6" = "American Indian or Alaska Native, Black or African American, Hispanic or Latinx, White",
            "1,4" = "American Indian or Alaska Native, Hispanic or Latinx",
            "1,6" = "American Indian or Alaska Native, White",
            "2,3,4,6" = "Asian, Black or African American, Hispanic or Latinx, White",
            "2,3,6" = "Asian,Black or African American, White",
            "2,6" = "Asian, White",
            "3,4" = "Black or African American, Hispanic or Latinx",
            "3,4,6" = "Black or African American, Hispanic or Latinx, White",
            "3,6" = "Black or African American, White",
            "4,6" = "Hispanic or Latinx, White"
              ))

leader_char <- leader_char %>% 
  mutate(leader1_race_agg = recode(leader1_race,
            "0" = "Prefer not to say",
            "1" = "American Indian or Alaska Native",
            "2" = "Asian",
            "3" = "Black or African American",
            "4" = "Hispanic or Latinx",
            "5" = "Native Hawaiian or Pacific Islander",
            "6" = "White",
            "7" = as.character(leader1_gender_self_identify_text)
              ))
```

```{r}
leader_char %>% 
  group_by(leader1_race) %>% 
  summarise(n = n()) %>% 
  datatable()
```

```{r}
# table(leader_char$leader1_race_agg)
```



# Q2: What are the racial/ethnic and gender identities of all Canopy school leaders? (How many schools are led by school leaders of different racial/ethnic identities?)


# Q3: Are BIPOC school leaders more likely to be found in certain types of schools (e.g. urban, charter, ...)?


# Q4: How representative are Canopy school leaders of the student populations they serve?
