---
title: "staff_student_diversity"
author: "Merly Klaas"
date: "2023-06-08"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
library(pacman)
p_load(here, rio, janitor, purrr, tidyverse,parameters, DT, nFactors, stringr, forcats)
#source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
region <- import(here("data", "us_census_bureau_regions_and_divisions.csv" )) %>% 
  clean_names()
transcend_cols = c(
  "#1A4C81",
  "#59C3B4",
  "#EF464B",
  "#ADE0EE"
)

transcend_grays = c(
  "#4D4D4F",
  "#9D9FA2",
  "#D1D3D4"
)

 theme_2 <- function() {
   theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black", face="bold"),
    plot.subtitle = element_text(family = "Sans", size = rel(0.8), face ="italic"),
    plot.background = element_blank(),
    panel.border = element_rect(colour = transcend_grays[1], 
                                fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8),
   axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x =element_blank(),
  )
    }
  
 
 theme_3 <- function() {
 theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black", face="bold"),
    plot.subtitle = element_text(family = "Sans", size = rel(0.8), face ="italic"),
    plot.background = element_rect(fill="transparent", color=NA),
    panel.background = element_rect(fill='transparent'),
    panel.border = element_rect(colour = transcend_grays[1], 
                               fill = NA),
    strip.text = element_text(size = rel(0.8)),
   plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8),
   axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x =element_blank()
  )}
```

```{r}

dat <- full %>% 
  select(school_id, 
         school_state, 
         leadership_diversity, 
         teaching_diversity, 
         pct_bipoc, 
         self_reported_locale_rural,
         self_reported_locale_suburban,
         self_reported_locale_urban,
         self_reported_locale_multiple) %>% 
  drop_na(pct_bipoc) %>% 
  filter(leadership_diversity != 0,
         teaching_diversity != 0) %>% 
  mutate(scale_bipoc = case_when(pct_bipoc<25 ~ 1,
                                 pct_bipoc>25 & pct_bipoc <50 ~ 2, 
                                 pct_bipoc >49 & pct_bipoc<75 ~ 3, 
                                 pct_bipoc >74 ~ 4, TRUE ~ NA), 
         lead_stu_div = scale_bipoc/leadership_diversity,
         teach_stu_div = scale_bipoc/teaching_diversity,
         lead_stu_prop = case_when(lead_stu_div < "1" ~ "Overly Representative",
                                   lead_stu_div == 1 ~ "Equally Representative",
                                   lead_stu_div > 1 & lead_stu_div < 3 ~ "Under Representative",
                                   teach_stu_div >=3  ~ "Severely under representative",
                                   TRUE ~ NA),
         teach_stu_prop = case_when(teach_stu_div < "1" ~ "Overly Representative",
                                   teach_stu_div == 1 ~ "Equally Representative",
                                   teach_stu_div > 1 & teach_stu_div < 3 ~ "Under Representative",
                                   teach_stu_div >=3  ~ "Severely under representative",
                                   TRUE ~ NA)) 
              
```

### Leadership & Teaching Team VS Student Diversity

In this figure, I wanted to show the ratio of Bipoc student compared to Bipoc teaching team / leadership team. Teaching team and Leadership team racial/ethnicity data were collected with four options: 1 = 0-24%, 2= 25-49%, 3= 50-74%, and  4= 75-100 % (I dropped NA and refuse to answer response). While we have the raw % of bipoc students. To have an equal comparison to leadership/teaching staff, I converted raw % of bipoc student into the same scale with four options (1-4). I then, created new variable of the scale of bipoc student on bipoc staff ratio by dividing bipoc student scale on bipoc staff. Lastly, I created another variable with four categories and created a bar chart showing #n of school within these categories: 

+ Severely under representative if ratio = 3:1 - 4:1
+ Under Representative if ratio 2:1
+ Equally Representative 1:1
+ Over Representative if staff team is more diverse"

```{r fig.height=7, fig.width=7}
dat %>% 
  drop_na(lead_stu_prop) %>% 
  ggplot(aes(x= factor(lead_stu_prop, levels = c("Severely under representative",
                                                 "Under Representative",
                                                 "Equally Representative",
                                                 "Overly Representative")), 
             fill = lead_stu_prop))+
  geom_bar(stat = "count", position = "dodge") + 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1, size = 3) + 
  labs(title = "How representative are leadership teams of the student populations they serve? \nRatio of BIPOC Students Compared to BIPOC Teaching Team",
       x="") +
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Severely under representative",
                              "Under Representative",
                              "Equally Representative",
                              "Overly Representative"),
  labels = c("Severely under representative \n(3:1 - 4:1)",
             "Under Representative \n(2:1)",
             "Equally Representative \n(1:1)",
             "Overly Representative \n(Leadership team is more diverse)"))+
  theme_2()

dat %>% 
  drop_na(teach_stu_prop) %>% 
  ggplot(aes(x= factor(teach_stu_prop, levels = c("Severely under representative",
                                                  "Under Representative",
                                                  "Equally Representative",
                                                  "Overly Representative")), 
             fill = teach_stu_prop))+
  geom_bar(stat = "count", position = "dodge") + 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1, size = 3) + 
  labs(title = "How representative are teaching teams of the student populations they serve? \nRatio of BIPOC Students Compared to BIPOC Teaching Team",
       x="") +
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Severely under representative",
                              "Under Representative",
                              "Equally Representative", 
                              "Overly Representative"),
                     labels = c("Severely under representative \n3:1 - 4:1",
                                "Under Representative \n2:1","Equally Representative \n1:1",
                                "Over Representative \nTeaching team is more diverse"))+
  theme_2()

```

### School Leader vs Student Diversity gap

I extracted leaders' race and recoded them into Bipoc/non-bipoc for each school as binary variable (coded 1 for Bipoc). Then, I calculated percentage of Bipoc leader within each school. I joined this data with percent bipoc student in each school. 

In the first graph, I calculated overall percentage of bipoc leaders and bipoc students, then visualized it as bar graph. 

```{r include=FALSE}
dat2 <- confidential %>% 
  select(school_id, ends_with("race")) %>% 
  pivot_longer(cols = 2:12,
               names_to = "leader",
               names_pattern = "leader(.*)_race",
                values_to = "race") %>% 
  drop_na() %>% 
  filter(race != 0) %>% 
  mutate(leader_bipoc = case_when(race == 6 ~ 0,
                                  TRUE ~ 1)) %>% 
  group_by(school_id) %>% 
  mutate( pct_leader_bipoc = round(mean(leader_bipoc)*100,0)) %>% 
  distinct(school_id, .keep_all=TRUE )

stud_bipoc <- full %>% 
  select(school_id, school_state, pct_bipoc)

dat2<- left_join(dat2, stud_bipoc) %>% 
  drop_na(pct_bipoc)
               
```




```{r}
dat2 %>% 
  pivot_longer(cols = c(pct_leader_bipoc, pct_bipoc),
               names_to = "category",
               values_to = "pct") %>% 
  group_by(category) %>% 
  summarise(percent_total = mean(pct)) %>% 
  ggplot(aes(x=category, y=percent_total, fill = category)) +
  geom_col(width = .5)+ 
  geom_text(aes(x=category, y=percent_total, label = paste(round(percent_total,0),"%")), vjust= -.1)+ 
  scale_fill_manual(name ="",
                    values = transcend_cols,
                    labels=c("% Non-white student","% Non-white Leader"))+ 
  labs(title = "School Leader & Student Diversity Gap (Canopy School Sample)",
            subtitle = "66% of all students identify as Non-white while only 42% of all leaders identify as Non-white") + 
  theme_2()


```

In the second graph, I grouped by state and calculated percent total of bipoc leader and student in each state. I visualized them as dumbbell plot to highlight the racial diversity gap between school leader and student. 


```{r fig.height=8, fig.width=8}
#gap <-
dat2 %>% 
  pivot_longer(cols = c(pct_leader_bipoc, pct_bipoc),
               names_to = "category",
               values_to = "pct") %>% 
  group_by(school_state, category) %>% 
  summarise(percent_total = mean(pct)) %>% 
  drop_na() %>% 
  ggplot()+ 
  geom_line(aes(x=percent_total, 
                y = fct_reorder(school_state,percent_total), 
                group = school_state),
                linetype= "dotted", 
            color = "gray") +
  geom_text(aes(x=percent_total, y=school_state, label = round(percent_total,0)),
               nudge_x = 2, 
            size = 3, 
            family = "sans") +
  geom_point(aes(x=percent_total, 
                 y = fct_reorder(school_state,percent_total),
                 color = category), 
             size=3, 
             show.legend = TRUE) +
  scale_color_manual(name ="",
                    values = transcend_cols,
                    breaks = c("pct_leader_bipoc","pct_bipoc"),
                    labels=c("% BIPOC Leader","% BIPOC Student"))+ 
  labs(title = "School Leader & Student Diveristy Gap (Canopy School Sample)") +
  theme_3()
   
# ggsave("finding-5.png", plot = gap, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```

### Mean comparison

```{r include=FALSE}
anova_dat <- dat %>% 
  left_join(dat2) %>% 
  mutate(self_reported_locale_rural = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_rural),
         self_reported_locale_urban = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_urban),
         self_reported_locale_suburban = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_suburban)) %>% #recode if select multiple, other locale = 0
pivot_longer(cols = starts_with("self_"),
               names_to = "locale", 
               values_to = "count",
               names_prefix= "self_reported_locale_") %>% 
  filter(count != 0) %>% 
 # mutate_at(vars(c("leadership_diversity", "teaching_diversity", "scale_bipoc", "pct_leader_bipoc", "pct_bipoc")), scale) %>% 
  pivot_longer(cols = c(leadership_diversity, teaching_diversity, scale_bipoc, pct_leader_bipoc, pct_bipoc),
               names_to = "category", 
               values_to = "pct")
  
  
```

### t-test: 
#### *Does the % of BIPOC leader is significantly different than the % of Bipoc Student?* 

Yes! Overall, the percentage of BIPOC leader is significantly lower than the % of Bipoc Student.
```{r}
t1 <- anova_dat %>% 
  filter(category == "pct_leader_bipoc" | category == "pct_bipoc")
  t.test(pct ~ category, data=t1)

t1 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()
#locale<-
t1 %>% 
  ggplot(aes(category, pct, fill=locale))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)
# ggsave("finding-6.png", plot = locale, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")

  
  
```

#### *Does the scaled number of BIPOCof BIPOC teaching team is significantly different than Bipoc Student?* 

**Yes! Overall, the percentage of BIPOC teaching team is significantly lower than the % of Bipoc Student.**

Remember that for teaching & leadership team diversity, we use 1 - 4, so does the scale_bipoc 

```{r}

t2 <- anova_dat %>% 
  filter(category == "teaching_diversity" | category == "scale_bipoc",
         pct != 5) #answer not sure)

t.test(pct ~ category, data=t2)



t2 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()


```

#### *Does the scaled number of BIPOC Leadership team is significantly different than  Bipoc Student?* 

**Yes! Overall, the scaled number of BIPOC Leadership team is significantly lower than the Bipoc Student.**
```{r}
t3 <- anova_dat %>% 
  filter(category == "leadership_diversity" | category == "scale_bipoc",
         pct != 5) #answer not sure)

t.test(pct ~ category, data=t3)


t3 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()

```

 
### Anova / F Test
 
```{r}

model1 <- aov(pct ~ category * locale, data = t1)
summary(model1)


```


