---
title: "staff_student_diversity"
author: "Merly Klaas"
date: "2023-06-08"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r include=FALSE}
library(pacman)
p_load(here, rio, janitor, purrr, tidyverse,parameters, DT, nFactors, stringr, forcats,ggpubr)
#source(here("scripts/branding.R"))
load(here("data/complete_canopy_2023.RData"))
region <- import(here("data", "us_census_bureau_regions_and_divisions.csv" )) %>% 
  clean_names()
transcend_cols = c(
  "#1A4C81",
  "#59C3B4",
  "#EF464B",
  "#ADE0EE"
)

transcend_grays = c(
  "#4D4D4F",
  "#9D9FA2",
  "#D1D3D4"
)

 theme_2 <- function() {
   theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black", face="bold"),
    plot.subtitle = element_text(family = "sans", size = rel(0.8)),
    plot.background = element_blank(),
     panel.background = element_rect(fill="transparent", color=NA),
    panel.border = element_rect(colour = transcend_grays[1], 
                                fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  #  legend.position = "bottom",
   legend.text=element_text(size=8),
   axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text =element_blank(),
  )
    }
  
 
 theme_3 <- function() {
 theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black"),
   plot.subtitle = element_text(family = "Sans", size = rel(0.8)),
    plot.background = element_rect(fill="transparent", color=NA),
    panel.background = element_rect(fill="transparent", color=NA),
    panel.border = element_rect(colour = transcend_grays[1], 
                            fill = NA),
    strip.text = element_text(size = rel(0.8)),
   plot.margin = margin(10, 24, 10, 10, "pt"),
   # panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8),
   axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x =element_blank()
  )}
```

```{r}

dat <- full %>% 
  select(school_id, 
         school_state, 
         leadership_diversity, 
         teaching_diversity, 
         pct_bipoc,
         self_reported_total_enrollment,
         self_reported_locale_rural,
         self_reported_locale_suburban,
         self_reported_locale_urban,
         self_reported_locale_multiple) %>% 
  drop_na(pct_bipoc) %>% 
  filter(leadership_diversity != 0 |leadership_diversity != 5,
         teaching_diversity != 0|teaching_diversity != 5) %>% 
  mutate(scale_bipoc = case_when(pct_bipoc<25 ~ 1,
                                 pct_bipoc>25 & pct_bipoc <50 ~ 2, 
                                 pct_bipoc >49 & pct_bipoc<75 ~ 3, 
                                 pct_bipoc >74 ~ 4, TRUE ~ NA), 
         lead_stu_div = scale_bipoc/leadership_diversity,
         teach_stu_div = scale_bipoc/teaching_diversity,
         lead_stu_prop = case_when(lead_stu_div < "1" ~ "Overly Representative",
                                   lead_stu_div == 1 ~ "Equally Representative",
                                   lead_stu_div > "1" ~ "Under Representative", 
                                 #  & lead_stu_div < 3 ,
                                 #  teach_stu_div >=3  ~ "Severely under representative",
                                   TRUE ~ NA),
         teach_stu_prop = case_when(teach_stu_div < "1" ~ "Overly Representative",
                                   teach_stu_div == 1 ~ "Equally Representative",
                                   teach_stu_div > 1 ~ "Under Representative",
                                   #& teach_stu_div < 3 ,
                              #     teach_stu_div >=3  ~ "Severely under representative",
                                   TRUE ~ NA)) 
              
```

### Leadership & Teaching Team VS Student Diversity

In this figure, I wanted to show the ratio of Bipoc student compared to Bipoc teaching team / leadership team. Teaching team and Leadership team racial/ethnicity data were collected with four options: 1 = 0-24%, 2= 25-49%, 3= 50-74%, and  4= 75-100 % (I dropped NA and refuse to answer response). While we have the raw % of bipoc students. To have an equal comparison to leadership/teaching staff, I converted raw % of bipoc student into the same scale with four options (1-4). I then, created new variable of the scale of bipoc student on bipoc staff ratio by dividing bipoc student scale on bipoc staff. Lastly, I created another variable with four categories and created a bar chart showing #n of school within these categories: 

+ Severely under representative if ratio = 3:1 - 4:1
+ Under Representative if ratio 2:1
+ Equally Representative 1:1
+ Over Representative if staff team is more diverse"

```{r eval=FALSE, include=FALSE}
leader <- dat %>% 
  drop_na(lead_stu_prop) %>% 
  ggplot(aes(x= factor(lead_stu_prop, levels = c("Severely under representative",
                                                 "Under Representative",
                                                 "Equally Representative",
                                                 "Overly Representative")), 
             fill = lead_stu_prop))+
  geom_bar(stat = "count", position = "dodge") + 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1, size = 3) + 
  labs(subtitle = "BIPOC Students Compared to BIPOC Leadership Team",
       x="") +
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Severely under representative",
                              "Under Representative",
                              "Equally Representative",
                              "Overly Representative"),
  labels = c("Severely under representative \n(3:1 - 4:1)",
             "Under Representative \n(2:1 or 3:2)",
             "Equally Representative \n(1:1)",
             "Overly Representative \n(School personel is more diverse)"))+
  theme_2()  +
  theme(legend.position = "none")

teaching <- dat %>% 
  drop_na(teach_stu_prop) %>% 
  ggplot(aes(x= factor(teach_stu_prop, levels = c("Severely under representative",
                                                  "Under Representative",
                                                  "Equally Representative",
                                                  "Overly Representative")), 
             fill = teach_stu_prop))+
  geom_bar(stat = "count", position = "dodge") + 
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -1, size = 3) + 
  labs(subtitle =   "BIPOC Students Compared to BIPOC Teaching Team",
       x="") +
   scale_fill_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Severely under representative",
                              "Under Representative",
                              "Equally Representative", 
                              "Overly Representative"),
                     labels=c("Severely under representative \n3:1 - 4:1",
                              "Under Representative \n(2:1 or 3:2)",
                              "Equally Representative \n1:1",
                               "Overly Representative \n(School personel is more diverse)"))+
 theme_2()+ 
  theme(legend.position = "none")



fig <- ggpubr::ggarrange(leader + teaching, common.legend = TRUE, legend="bottom")

diversity <- annotate_figure(fig, top = text_grob("How representative are the school personel of the student populations they serve?",family = "Bebas Neue", color = "black", size = 14))

#ggsave("finding-5-V2.png", plot = diversity, path = here("final_products", "draft-findings"),
   #   width = 8, height = 6, units = "in")
```

### Scatter plot for Teaching Diversity

```{r}
sct_teach <- dat %>% 
  select(teaching_diversity,scale_bipoc, teach_stu_prop) %>% 
  drop_na() %>% 
  filter(teaching_diversity != 5) %>% 
  ggplot(aes(scale_bipoc, teaching_diversity)) + 
  geom_jitter(aes(color=teach_stu_prop)) +
   geom_abline(intercept = 0,
              slope = 1, 
              linetype = "dashed",
              size = 0.5,
              color = transcend_grays[1])+
  labs(title = "Teaching Team & Student Diversity Gap",
       subtitle = "Majority of School has lower number of BIPOC teaching team compared to BIPOC Student",
       x= "BIPOC Student",
       y="BIPOC Teaching Team",
       caption = "Note: Each dot represents one school")+
  scale_color_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Overly Representative",
                             "Equally Representative",
                             "Under Representative")) +
      
   theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.subtitle = element_text(family = "sans", size = rel(0.8)),
    plot.background = element_rect(fill="transparent", color=NA),
   panel.background = element_rect(fill="transparent", color=NA),
    panel.border = element_rect(colour = transcend_grays[1], 
                                fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
   # panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8)
   #axis.title = element_blank()
   # axis.ticks = element_blank(),
   # axis.text =element_blank(),
  )
sct_teach
#ggsave("finding-5-V3.png", plot = sct_teach, path = here("final_products", "draft-findings"),
 #     width = 8, height = 6, units = "in")

   
```

### Scatter plot for Leadership Diversity
```{r}

sct_lead <- dat %>% 
  select(leadership_diversity,scale_bipoc, lead_stu_prop) %>% 
  drop_na() %>% 
  filter(leadership_diversity != 5) %>% 
  ggplot(aes(scale_bipoc, leadership_diversity)) + 
  geom_jitter(aes(color=lead_stu_prop)) +
   geom_abline(intercept = 0,
              slope = 1, 
              linetype = "dashed",
              size = .5,
              color = transcend_grays[1])+
  labs(title = "Leadership Team & Student Diversity Gap",
       subtitle = "Majority of School has lower number of BIPOC Leadership team compared to BIPOC Student",
       x= "BIPOC Student",
       y="BIPOC Leadership Team",
       caption = "Note: Each dot represents one school")+
  scale_color_manual(name= "",
                     values = transcend_cols,
                     breaks=c("Overly Representative",
                             "Equally Representative",
                             "Under Representative")) +
      
   theme(
     plot.title.position = "plot",
     plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.subtitle = element_text(family = "sans", size = rel(0.8)),
    plot.background = element_rect(fill="transparent", color=NA),
   panel.background = element_rect(fill="transparent", color=NA),
    panel.border = element_rect(colour = transcend_grays[1], 
                                fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
   # panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
   legend.text=element_text(size=8)
   #axis.title = element_blank()
   # axis.ticks = element_blank(),
   # axis.text =element_blank(),
  )
sct_lead

#ggsave("finding-5b-V3.png", plot = sct_lead, path = here("final_products", "draft-findings"),
   #   width = 8, height = 6, units = "in")

```




### School Leader vs Student Diversity gap

I extracted leaders' race and recoded them into Bipoc/non-bipoc for each school as binary variable (coded 1 for Bipoc). Then, I calculated percentage of Bipoc leader within each school. I joined this data with percent bipoc student in each school. 

In the first graph, I calculated overall percentage of bipoc leaders and bipoc students, then visualized it as bar graph. 

```{r include=FALSE}
dat2 <- confidential %>% 
  select(school_id, ends_with("race")) %>% 
  pivot_longer(cols = 2:12,
               names_to = "leader",
               names_pattern = "leader(.*)_race",
                values_to = "race") %>% 
  drop_na() %>% 
  filter(race != 0) %>% 
  mutate(leader_bipoc = case_when(race == 6 ~ 0,
                                  TRUE ~ 1)) %>% 
  group_by(school_id) %>% 
  mutate( pct_leader_bipoc = round(mean(leader_bipoc)*100,0)) %>% 
  distinct(school_id, .keep_all=TRUE )

stud_bipoc <- full %>% 
  select(school_id, school_state, pct_bipoc)

dat2<- left_join(dat2, stud_bipoc) %>% 
  drop_na(pct_bipoc)
               
```




```{r}
dat2 %>% 
  pivot_longer(cols = c(pct_leader_bipoc, pct_bipoc),
               names_to = "category",
               values_to = "pct") %>% 
  group_by(category) %>% 
  summarise(percent_total = mean(pct)) %>% 
  ggplot(aes(x=category, y=percent_total, fill = category)) +
  geom_col(width = .5)+ 
  geom_text(aes(x=category, y=percent_total, label = paste(round(percent_total,0),"%")), vjust= -.1)+ 
  scale_fill_manual(name ="",
                    values = transcend_cols,
                    labels=c("% Non-white student","% Non-white Leader"))+ 
  labs(title = "School Leader & Student Diversity Gap (Canopy School Sample)",
            subtitle = "66% of all students identify as Non-white while only 42% of all leaders identify as Non-white") + 
  theme_2()


```

In the second graph, I grouped by state and calculated percent total of bipoc leader and student in each state. I visualized them as dumbbell plot to highlight the racial diversity gap between school leader and student. 


```{r fig.height=8, fig.width=8}
#gap <-
dat2 %>% 
  pivot_longer(cols = c(pct_leader_bipoc, pct_bipoc),
               names_to = "category",
               values_to = "pct") %>% 
  group_by(school_state, category) %>% 
  summarise(percent_total = mean(pct)) %>% 
  drop_na() %>% 
  ggplot()+ 
  geom_line(aes(x=percent_total, 
                y = fct_reorder(school_state,percent_total), 
                group = school_state),
                linetype= "dotted", 
            color = "gray") +
  geom_text(aes(x=percent_total, y=school_state, label = round(percent_total,0)),
               nudge_x = 2, 
            size = 3, 
            family = "sans") +
  geom_point(aes(x=percent_total, 
                 y = fct_reorder(school_state,percent_total),
                 color = category), 
             size=3, 
             show.legend = TRUE) +
  scale_color_manual(name ="",
                    values = transcend_cols,
                    breaks = c("pct_leader_bipoc","pct_bipoc"),
                    labels=c("% BIPOC Leader","% BIPOC Student"))+ 
  labs(title = "School Leader & Student Diveristy Gap (Canopy School Sample)") +
  theme_3()
   
# ggsave("finding-5.png", plot = gap, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")
```


school leader and student Gap by State
```{r fig.height=8, fig.width=8}
#gap <-
dat2 %>% 
  pivot_longer(cols = c(pct_leader_bipoc, pct_bipoc),
               names_to = "category",
               values_to = "pct") %>% 
  group_by(school_state, category) %>% 
  summarise(percent_total = mean(pct)) %>% 
  drop_na() %>% 
  ggplot()+ 
  geom_line(aes(x=percent_total, 
                y = fct_reorder(school_state,percent_total), 
                group = school_state),
                linetype= "dotted", 
            color = "gray") +
  geom_text(aes(x=percent_total, y=school_state, label = round(percent_total,0)),
               nudge_x = 2, 
            size = 3, 
            family = "sans") +
  geom_point(aes(x=percent_total, 
                 y = fct_reorder(school_state,percent_total),
                 color = category), 
             size=3, 
             show.legend = TRUE) +
  scale_color_manual(name ="",
                    values = transcend_cols,
                    breaks = c("pct_leader_bipoc","pct_bipoc"),
                    labels=c("% BIPOC Leader","% BIPOC Student"))+ 
  labs(title = "School Leader & Student Diversity Gap (Canopy School Sample)") +
  theme_3()
   
```


### Mean comparison

```{r include=FALSE}
anova_dat <- dat %>% 
  left_join(dat2) %>% 
  mutate(self_reported_locale_rural = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_rural),
         self_reported_locale_urban = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_urban),
         self_reported_locale_suburban = case_when(self_reported_locale_multiple == "1" ~ 0, TRUE ~ self_reported_locale_suburban)) %>% #recode if select multiple, other locale = 0
pivot_longer(cols = starts_with("self_reported_locale_"),
               names_to = "locale", 
               values_to = "count",
               names_prefix= "self_reported_locale_") %>% 
  filter(count != 0) %>% 
 # mutate_at(vars(c("leadership_diversity", "teaching_diversity", "scale_bipoc", "pct_leader_bipoc", "pct_bipoc")), scale) %>% 
  pivot_longer(cols = c(leadership_diversity, teaching_diversity, scale_bipoc, pct_leader_bipoc, pct_bipoc),
               names_to = "category", 
               values_to = "pct")
  
  
```
### Dumbbell school leader and student Gap by Locale
```{r fig.height=4, fig.width=6}
locale <- 
  anova_dat %>% 
 filter(category == "pct_leader_bipoc" | category == "pct_bipoc") %>% 
  mutate(locale = str_to_title(locale)) %>% 
  group_by(locale, category) %>% 
  summarise(percent_total = weighted.mean(pct,self_reported_total_enrollment, na.rm = TRUE)) %>% 
  drop_na() %>%
  pivot_wider(names_from = category,
              values_from = percent_total) %>% 
  mutate(gap = paste(round(pct_bipoc - pct_leader_bipoc,0),"%")) %>% 
  pivot_longer(cols = c(pct_bipoc, pct_leader_bipoc),
               names_to = "category",
               values_to = "percent_total") %>% 
  ggplot()+ 
  geom_line(aes(x=percent_total, 
                y = fct_reorder(locale,percent_total), 
                group = locale),
                linetype= "dotted", 
            color = "gray") +
  geom_text(aes(x=percent_total, y=locale,
                label = paste(round(percent_total,0),"%")),
               nudge_x = -3, 
            size = 3, 
            family = "sans",
            fontface="bold") +
  geom_point(aes(x=percent_total, 
                 y = fct_reorder(locale,percent_total),
                 color = category), 
             size=4, 
             show.legend = TRUE) +
    scale_x_continuous(expand = c(0,0),
                     limits = c(0,100)) + 
  coord_equal(ratio = 7)+
  scale_color_manual(name ="",
                    values = transcend_cols,
                    breaks = c("pct_leader_bipoc","pct_bipoc"),
                    labels=c("% BIPOC Leader","% BIPOC Student"))+ 
  labs(title = "School Leader & Student Diversity Gap",
      subtitle ="Overall, 66% of all students identify as BIPOC while only 42% of all school leaders identify as BIPOC") +
   theme(
     plot.title.position = "plot",
    plot.title = element_text(family = "Bebas Neue", color = "black"),
    plot.subtitle = element_text(family = "Sans", size = rel(0.8)),
    plot.background = element_blank(),
    panel.grid.major.x = element_line(color="gray", size= 0.1),
    panel.background = element_blank(),
    axis.title = element_blank(),
   panel.border = element_rect(colour = transcend_grays[1], 
                                fill = NA),
    strip.text = element_text(size = rel(0.8)),
    plot.margin = margin(10, 24, 10, 10, "pt"),
    panel.grid.major.y= element_blank(),
    legend.position = "bottom"
  ) 
locale
```





```{r}
# a data frame with all the annotation info
annotation <- data.frame(
   x = c(60,45,35,20),
   y = c(4.3,3.3,2.3,1.3),
   label = c("Gap= 28%", "Gap= 41%", 
             "Gap= 17%", "Gap= 23%")
)

fig <- locale + geom_label(data=annotation, aes( x=x, y=y, label=label),     
           color=transcend_cols[3], 
           size=2.5 , angle=45, fontface="bold")

fig
#ggsave("finding-6-V3.png", plot = fig, path = here("final_products", "draft-findings"),
  #    width = 8, height = 6, units = "in")


```


### t-test: 
#### *Does the % of BIPOC leader is significantly different than the % of Bipoc Student?* 

Yes! Overall, the percentage of BIPOC leader is significantly lower than the % of Bipoc Student.
```{r}
t1 <- anova_dat %>% 
  filter(category == "pct_leader_bipoc" | category == "pct_bipoc")
  t.test(pct ~ category, data=t1)

t1 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()
#locale<-
t1 %>% 
  ggplot(aes(category, pct, fill=locale))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)
# ggsave("finding-6.png", plot = locale, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")

  
  
```

#### *Does the scaled number of BIPOCof BIPOC teaching team is significantly different than Bipoc Student?* 

**Yes! Overall, the percentage of BIPOC teaching team is significantly lower than the % of Bipoc Student.**

Remember that for teaching & leadership team diversity, we use 1 - 4, so does the scale_bipoc 

```{r}

t2 <- anova_dat %>% 
  filter(category == "teaching_diversity" | category == "scale_bipoc",
         pct != 5) #answer not sure)

t.test(pct ~ category, data=t2)



t2 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()


```

#### *Does the scaled number of BIPOC Leadership team is significantly different than  Bipoc Student?* 

**Yes! Overall, the scaled number of BIPOC Leadership team is significantly lower than the Bipoc Student.**
```{r}
t3 <- anova_dat %>% 
  filter(category == "leadership_diversity" | category == "scale_bipoc",
         pct != 5) #answer not sure)

t.test(pct ~ category, data=t3)


t3 %>% 
  ggplot(aes(category, pct, fill=category))+
  geom_boxplot()+
  scale_fill_manual(values = transcend_cols)+ 
  theme_2()

```

 
### Anova / F Test
 
```{r}

model1 <- aov(pct ~ category * locale, data = t1)
summary(model1)
model2 <- aov(pct ~ category * locale, data = t2)
summary(model2)
model3 <- aov(pct ~ category * locale, data = t3)
summary(model3)

```


