---
title: "Updated Tagging Patterns"
author: "Karen Zyskind"
date: "2023-05-16"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(reshape2)
library(cowplot)
library(dplyr)
library("pacman")
library("DT")
pacman::p_load("summarytools")
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2023.RData"))
```


# How do tagging patterns differ across states?
First re-arranging the data 

```{r State tag df, echo=FALSE}
# Load required libraries

# Extracted the columns representing the tags
tag_columns <- colnames(tags)[-1]  # Exclude the first column (school_id)

# Reshaped the data to long format
tags_long <- tags %>%
  tidyr::pivot_longer(cols = tag_columns, names_to = "Tag", values_to = "Count") %>%
  filter(Count > 0)  # Remove rows with zero count (if applicable)

# Merge with the schools dataframe
merged_data <- merge(tags_long, schools, by = "school_id")

# Calculated summary statistics by school_state and tag
state_tag_summary <- merged_data %>%
  group_by(school_state, Tag) %>%
  summarize(
    Total_Count = sum(Count),
    Unique_Schools = n_distinct(school_id)
  ) %>%
  arrange(desc(Total_Count))

# View the summary statistics
print(state_tag_summary)

print(state_tag_summary)
print(merged_data)

library(ggplot2)
library(dplyr)
library(stringr)

str(state_tag_summary)
```

## Now finding the distributions 
Top states include 
1. California with 1252 different tags 
2 Colorado with 1014
3 Massachusetts with 903
4 New York with 791
5 Arizona with 641
```{r state,echo=FALSE}

library(dplyr)
library(knitr)

# Overall distribution of tags
tag_summary <- state_tag_summary %>%
  group_by(Tag) %>%
  summarize(Total_Count = sum(Total_Count), .groups = "drop") %>%
  arrange(desc(Total_Count))

kable(tag_summary, caption = "Overall Distribution of Tags")

# Exploring distribution of tags within each state
state_summary <- state_tag_summary %>%
  group_by(school_state) %>%
  summarize(Total_Count = sum(Total_Count), .groups = "drop") %>%
  arrange(desc(Total_Count))

kable(state_summary, caption = "Distribution of Tags Within Each State")

# Comparing the top tags across states
top_states <- head(state_summary, 5)  # Select top 5 states with highest tag counts

filtered_data <- state_tag_summary %>%
  filter(school_state %in% top_states$school_state)

state_tag_comparison <- filtered_data %>%
  group_by(school_state, Tag) %>%
  summarize(Total_Count = sum(Total_Count), .groups = "drop") %>%
  arrange(school_state, desc(Total_Count))

kable(state_tag_comparison, caption = "Comparison of Tags Across States")



```

## Including Plots for tags across states 
**Been having issues with font**


```{r chart, echo=FALSE}
# Set the font size and label width for the x-axis labels
label_size <- 10
label_width <- 15

# Create the bar chart
bar_chart <- ggplot(state_tag_summary, aes(x = reorder(school_state, -Total_Count), y = Total_Count)) +
  geom_bar(stat = "identity", fill = transcend_cols[1]) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = label_width)) +
  labs(x = "State", y = "Total Count of Tags", title = "Tag Counts by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = label_size, family = "Arial"),
    axis.text.y = element_text(size = label_size, family = "Arial"),
    plot.title = element_text(family = "Arial")
  )

# Display the bar chart
print(bar_chart)
# ggsave("finding-12b.png", plot = bar_chart, path = here("final_products", "draft-findings"),
#        width = 12, height = 8, units = "in")

```

## Type and Pattern of Tags 
First looking at type count 
Charter n= 113
District n= 102
Independent = 36
```{r Type, echo=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)

# Load the branding file
source(here("scripts", "branding.R"))

# Count the number of schools for each school descriptor
school_counts <- schools %>%
  count(school_descriptor)

# Create the table
school_counts_table <- school_counts %>%
  mutate(school_descriptor = c("Charter", "District", "Independent")) %>%
  kable(caption = "Number of Schools by Descriptor", align = "c") %>%
  kable_styling("striped")

# Display the table
print(school_counts_table)


```

# Differences of tags across Type 
```{r Differences, echo=FALSE}

library(dplyr)
library(knitr)
library(kableExtra)

# Merge the tags_long and schools data frames
merged_data <- inner_join(tags_long, schools, by = "school_id")

# Define the custom labels for school_descriptor
school_descriptor_labels <- c("Public district school", "Public charter school", "Independent (private) school")

# Filter necessary columns for the analysis and apply custom labels
filtered_data <- merged_data %>%
  select(Tag, school_descriptor) %>%
  mutate(school_descriptor = case_when(
    school_descriptor == 1 ~ school_descriptor_labels[1],
    school_descriptor == 2 ~ school_descriptor_labels[2],
    school_descriptor == 3 ~ school_descriptor_labels[3],
    TRUE ~ as.character(school_descriptor)  # Handle other values if present
  ))

# Calculate differences in tags by school_descriptor (modify this step based on your desired calculation method)

# Create the table
tag_differences_table <- filtered_data %>%
  group_by(Tag, school_descriptor) %>%
  summarize(Total_Count = n()) %>%
  pivot_wider(names_from = school_descriptor, values_from = Total_Count, values_fill = 0) %>%
  kable(caption = "Differences in Tags by School Descriptor", align = "c") %>%
  kable_styling("striped")

# Display the table
print(tag_differences_table)




```
## Biggest Difference 
Will be calculating the absolute differnces between the counts of each count for pupblic and private
 Table of Differences 
The table shows the values for each tag, the counts in the public district school, independent (private) school, and public charter school, as well as the absolute differences between these counts.
```{r Tables Diff, echo=FALSE}

# Calculated counts for each tag and school descriptor
type_tbl <- filtered_data %>%
  group_by(Tag, school_descriptor) %>%
  summarize(Total_Count = n()) %>%
  pivot_wider(names_from = school_descriptor, values_from = Total_Count, values_fill = 0)

# Apply labels to school descriptor columns
colnames(type_tbl)[colnames(type_tbl) == "1"] <- "Public district school"
colnames(type_tbl)[colnames(type_tbl) == "2"] <- "Public charter school"
colnames(type_tbl)[colnames(type_tbl) == "3"] <- "Independent (private) school"

# Calculated the absolute differences between counts for each tag
type_tbl <- type_tbl %>%
  mutate(dis_char_diff = abs(`Public district school` - `Public charter school`),
         char_ind_diff = abs(`Public charter school` - `Independent (private) school`),
         ind_dis_diff = abs(`Independent (private) school` - `Public district school`))

# Display the table of differences
datatable(type_tbl)




```
##Differences from one another 


```{r}
str(type_tbl)
# Select the desired columns for the table
diff_table <- type_tbl %>%
  select(Tag, `Public district school`, `Independent (private) school`, `Public charter school`,
         dis_char_diff, char_ind_diff, ind_dis_diff)

# Print the table
print(diff_table)

dis_char_viz <- diff_table %>%
  slice_max(dis_char_diff, n = 10, with_ties = FALSE) %>%
  ggplot(aes(reorder(Tag, dis_char_diff), dis_char_diff, fill = Tag)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  geom_text(aes(x = Tag, y = dis_char_diff + 0.1, label = dis_char_diff),
            color = "black", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_y_continuous(limits = c(0, max(diff_table$dis_char_diff) + 1)) +
  labs(x = "",
       y = "Absolute Difference",
       title = "Top 10 Tags with Largest Differences between District and Charter Schools",
       subtitle = "Comparison within school types")

dis_char_viz



```

