---
title: "State Analysis"
author: "Anwesha Guha"
date: '2023-04-23'
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(here, rio, tidyverse, DT, janitor)

dat <- import(here("data", "raw-ip.csv"))
```

What patterns do we see in the practices that make up different schools' innovative models?

# Setup

Let's see what variables we are working with.
```{r}
dat_names <- tibble(variable.names(dat))
```

# STATE: How do tagging patterns differ across states?

```{r}
state_tags <- dat %>% 
  select(school_state, 20:92)
```

Looks like a few schools have missing information for the state that they're in. I am removing that variable for now, but we should come back and fill in that info when possible.

```{r}
state_tags <- state_tags %>% 
  
  mutate(as.character(starts_with("practices")))

state_tags <- state_tags %>% 
  filter(!is.na(school_state)) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  group_by(school_state, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE))
```


