---
title: "Tony_week1"
author: "Tony Daza"
date: "2023-04-28"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Week 1 Reflections   

**Week 1 Reflection**    

Week 1 was a little hectic as I was asked to prepared a last minute presentation for a conference so in truth, I did not have a chance to dive into the data as much as I would have liked.  

**Questions about the data**   

I am not entirely sure if I should be using the confidential-dataset, the public-dataset, or data from the data cleaning Rmd. 

I am struggling with interpreting the race data in the internal data set for school leaders. 


**Specific findings**   

One finding is that I need to figure out if schools really have zero percent BIPOC students or if it is a reporting issue. There seem to be a lot of schools that report zero percent for the different racial groups. 



```{r import, include=FALSE}
library(tidyverse)
library(rio)
library(here)
library(reshape2)
library(cowplot)


dat_pub <- import(here("data/public-facing-dataset.csv"))

dat_con <- import(here("data/confidential-dataset.csv"))


```


# Student Demographic Data  
## Student Race/Ethnicity  

### Frequency Graphs  

```{r demog}

summary(dat_con[157:166])


hist(dat_con$self_reported_race_white, col = "cornflowerblue", 
     main = paste("Histogram of Percent White"),
     xlab = "Percent White", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)


hist(dat_con$self_reported_race_black , col = "darkorchid2", 
     main = paste("Histogram of Percent Black"),
     xlab = "Percent Black", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)

hist(dat_con$self_reported_race_hispanic , col = "chocolate2", 
     main = paste("Histogram of Percent Hispanic"),
     xlab = "Percent Hispanic", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)

hist(dat_con$self_reported_race_asian , col = "seagreen2", 
     main = paste("Histogram of Percent Asian"),
     xlab = "Percent Asian", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)

hist(dat_con$self_reported_race_aian , col = "royalblue2", 
     main = paste("Histogram of Percent Native American"),
     xlab = "Percent Native American", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)

hist(dat_con$self_reported_race_multiple , col = "orangered1", 
     main = paste("Histogram of Percent Multiple Races"),
     xlab = "Percent Multiple Races", 
     ylab = "Number of Schools",
     axes = TRUE, plot = TRUE, labels = FALSE)



```


**It seems like the schools reporting zero percent of each ethnicity is larger than expected for all groups**


```{r barplot}

barplot(table(dat_con$self_reported_race_black), col = "darkorchid2", 
        main = "Percent of Black Students",
        xlab = "Percent Black Students",
        ylab = "Number of Schools")


```

Same information, different format. I think the barplot potentially shows the range better.


**Last year they used violin plots**

```{r violin plot}

ggplot(data = dat_con, aes(x = "", y = self_reported_race_black, fill = self_reported_race_black)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = 21) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "", y = "Number of Schools") +
  theme_classic() +
  theme(legend.position = "none")


```

I am not super familiar with violin plots so I need to revise this. I want all the schools to be in one plot though.

```{r violin data, eval=FALSE, include=FALSE}
# Selecting just the student race data and school name
dat_con_short <- dat_con[c(3,157:162,164)]

vio_blk <- ggplot(data = dat_con_short, aes(x = "", y = self_reported_race_black, fill = self_reported_race_black)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = 21) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "", y = "Number of Schools") +
  theme_classic() +
  theme(legend.position = "none")

vio_asian <- ggplot(data = dat_con, aes(x = "", y = self_reported_race_black, fill = self_reported_race_black)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = 21) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "", y = "Number of Schools") +
  theme_classic() +
  theme(legend.position = "none")


plot_grid(vio_race + facet_wrap(~Race, scales = "free_y"), ncol = 1)

# Realized there was a better way.
```

### Violin Plots  

**This should show all of the violin plots together**  

It might make sense to collapse non-white into one group for a visualization as well.  

I don't like how the scales are different but it is a work in progress.  

```{r orchestra}

# Selecting just the student race data and school name
dat_con_short <- dat_con[c(3,157:162,164)]

# Creating a long version of the data
dat_con_long <- melt(dat_con_short, id.vars = "school_name", 
                    variable.name = "Race", 
                    value.name = "Percentage")

# Removing the prefix in front of the race data
dat_con_long$Race <- gsub("^self_reported_race_", "", dat_con_long$Race)


# Violin plots, but it is by school so it needs to be merged to clean it.
 vio_race <- ggplot(data = dat_con_long, aes(x = " ", y = Percentage, fill = Race)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, fill = "white", outlier.shape = NA) +
  labs(x = "School", y = "Percentage", fill = "Race") +
  theme_classic()

plot_grid(vio_race + facet_wrap(~Race, scales = "free_y"), ncol = 1)



```

# School Leader Demographics   
## School Leader Race Data   

**This is a work in progress**
I am struggling with interpreting the race data in the internal data set. 

