---
title: "Tagging Patterns_states"
author: "Karen Zyskind"
date: "2023-05-11"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(reshape2)
library(cowplot)
library(dplyr)
library("pacman")
library("DT")
pacman::p_load("summarytools")
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2023.RData"))
```

#How do tagging patterns differ across states?
## What would this look like? 


```{r patterns, echo=FALSE}
# Load required libraries

# Extracted the columns representing the tags
tag_columns <- colnames(tags)[-1]  # Exclude the first column (school_id)

# Reshaped the data to long format
tags_long <- tags %>%
  tidyr::pivot_longer(cols = tag_columns, names_to = "Tag", values_to = "Count") %>%
  filter(Count > 0)  # Remove rows with zero count (if applicable)

# Merge with the schools dataframe
merged_data <- merge(tags_long, schools, by = "school_id")

# Calculate summary statistics by school_state and tag
state_tag_summary <- merged_data %>%
  group_by(school_state, Tag) %>%
  summarize(
    Total_Count = sum(Count),
    Unique_Schools = n_distinct(school_id)
  ) %>%
  arrange(desc(Total_Count))

# View the summary statistics
print(state_tag_summary)

print(state_tag_summary)
print(merged_data)

library(ggplot2)
library(dplyr)
library(stringr)

# Create a bar chart
bar_chart <- ggplot(state_tag_summary, aes(x = reorder(school_state, -Total_Count), y = Total_Count)) +
  geom_bar(stat = "identity", fill = transcend_cols[1]) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  labs(x = "State", y = "Total Count of Tags", title = "Tag Counts by State") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, family = "Arial"),
    axis.text.y = element_text(family = "Arial"),
    plot.title = element_text(family = "Arial")
  )

# Display the bar chart
print(bar_chart)



```


```{r}
library(ggplot2)
library(treemapify)

# Create a treemap
treemap <- ggplot(state_tag_summary, aes(area = Total_Count, fill = school_state, label = school_state)) +
  geom_treemap() +
  geom_treemap_text(size = 5, color = "white") +
  labs(title = "Tagging Patterns Across States (Treemap)") +
  theme_minimal() +
  theme(panel.grid = element_blank())

# Display the treemap
print(treemap)


```

```{r}
library(ggplot2)

# Create the box plot
box_plot <- ggplot(state_tag_summary, aes(x = school_state, y = Total_Count)) +
  geom_boxplot(fill = "blue") +
  labs(x = "State", y = "Total Count of Tags", title = "Tagging Patterns Across States") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the box plot
print(box_plot)
#Each box represents the distribution of the Total_Count variable for a state. The x-axis represents the states, and the y-axis represents the total count of tags.



```
# How do tagging patterns differ among charter, district, and independent schools?

The difference in standard deviation between "Independent (private) school" and "Public charter school" is approximately 8.91.
The difference in standard deviation between "Independent (private) school" and "Public district school" is approximately 23.1.
The difference in standard deviation between "Public charter school" and "Public district school" is approximately 23.5

The differences in standard deviations between the different types of schools refer to the use of practices. The pivot table summarizes the total counts of practices for each school type. The standard deviations are then calculated for each school type, indicating the variability or spread of practice counts within each school type.

`



```

