---
title: "Differences Type"
author: "Karen Zyskind"
date: "2023-05-31"
html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
load("~/Desktop/canopy-dac-23/data/complete_canopy_2023.RData")
library(tidyverse)
library(here)
library(reshape2)
library(cowplot)
library(dplyr)
library("pacman")
library("DT")
pacman::p_load("summarytools")
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2023.RData"))

```

#Guiding questions/tasks & format:  
`SCHOOL TYPE`: How do tagging patterns differ among district, charter, independent, sws, virtual, microschool, homeschool, and hub schools?  
* table of biggest differences between categories  

##Type and how many tags
First recall Type and Pattern of Tags 
First looking at type count 
Charter n= 113
District n= 102
Independent = 36

```{r Type, echo=FALSE}
# Count the number of schools for each school descriptor
school_counts <- schools %>%
  count(school_descriptor)

# Load the necessary packages
# library(gt)
# Karen: I used datatable() from {DT} instead because I was having issues with {gt}

# Create the table
school_counts_table <- school_counts %>%
  mutate(school_descriptor = c("Charter", "District", "Independent")) #%>%
  gt() %>%
  tab_header(
    title = "Number of Schools by Descriptor"
  ) %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = cells_body(columns = everything())
  )

# Display the table
school_counts_table


```

## Counts of practices  by type first 
```{r countandtype, echo=FALSE}

library(dplyr)
library(tidyverse)
library(htmlTable)

# Subset relevant columns from the 'tags' data frame
practice_columns <- grep("^practices", names(tags), value = TRUE)

# Merge 'tags' and 'schools' data frames
merged_data <- merge(tags, schools, by = "school_id")

# Group the data by school descriptor and calculate practice counts
practice_counts <- merged_data %>%
  group_by(school_descriptor) %>%
  summarise(across(all_of(practice_columns), sum))

# Rename school descriptor values with labels
practice_counts$school_descriptor <- recode(practice_counts$school_descriptor,
                                            `1` = "Public district school",
                                            `2` = "Public charter school",
                                            `3` = "Independent")

# Arrange the table by school descriptor
practice_counts <- practice_counts %>%
  arrange(school_descriptor)

# Print the table
knitr::kable(practice_counts)

# Create a pivot table
pivot_table <- practice_counts %>%
  pivot_longer(cols = starts_with("practices_"), names_to = "practice", values_to = "count")

# Display the pivot table as an HTML table
htmlTable(pivot_table, rnames = FALSE)


```



## Differences table by percents for each practice by each type 
District vs Charter Difference: The percentage difference in tag occurrence between public district schools and public charter schools. Positive values indicate higher occurrence in district schools, while negative values indicate higher occurrence in charter schools.

```{r Differences, echo=FALSE}
library(DT)

# Create a vector to map numeric values to categories
descriptor_categories <- c("Public district school", "Public charter school", "Independent")

# Convert school_descriptor column to factor with corresponding categories
school_categories <- factor(schools$school_descriptor, levels = 1:3, labels = descriptor_categories)

# Filter the columns in the tags dataframe that start with "practice"
practice_cols <- grep("^practice", names(tags), value = TRUE)

# Create an empty dataframe to store the results
table_of_differences <- data.frame(
  Tag = character(),
  `District vs Charter Difference` = numeric(),
  `Charter vs Independent Difference` = numeric(),
  `Independent vs District Difference` = numeric(),
  `Pct Tags Within(District)` = numeric(),
  `Pct Tags Within(Charter)` = numeric(),
  `Pct Tags Within(Independent)` = numeric(),
  `Pct Tags Across(District)` = numeric(),
  `Pct Tags Across(Charter)` = numeric(),
  `Pct Tags Across(Independent)` = numeric(),
  `N Tags(District)` = numeric(),
  `N Tags(Charter)` = numeric(),
  `N Tags(Independent)` = numeric(),
  stringsAsFactors = FALSE
)

# Iterate over each tag column and calculate the differences and percentages
for (col in practice_cols) {
  tag <- col  # Assign the column name to the 'tag' variable as is
  
  district_pct <- round(sum(tags[, col] & school_categories == "Public district school") / sum(school_categories == "Public district school") * 100, 2)
  charter_pct <- round(sum(tags[, col] & school_categories == "Public charter school") / sum(school_categories == "Public charter school") * 100, 2)
  independent_pct <- round(sum(tags[, col] & school_categories == "Independent") / sum(school_categories == "Independent") * 100, 2)
  
  district_vs_charter_diff <- round(district_pct - charter_pct, 2)
  charter_vs_independent_diff <- round(charter_pct - independent_pct, 2)
  independent_vs_district_diff <- round(independent_pct - district_pct, 2)
  
  district_pct_within <- round(sum(tags[, col] & school_categories == "Public district school") / sum(tags[, col]) * 100, 2)
  charter_pct_within <- round(sum(tags[, col] & school_categories == "Public charter school") / sum(tags[, col]) * 100, 2)
  independent_pct_within <- round(sum(tags[, col] & school_categories == "Independent") / sum(tags[, col]) * 100, 2)
  
  district_pct_across <- round(sum(tags[, col] & school_categories != "Public district school") / sum(tags[, col]) * 100, 2)
  charter_pct_across <- round(sum(tags[, col] & school_categories != "Public charter school") / sum(tags[, col]) * 100, 2)
  independent_pct_across <- round(sum(tags[, col] & school_categories != "Independent") / sum(tags[, col]) * 100, 2)
  
  n_tags_district <- sum(tags[, col] & school_categories == "Public district school")
  n_tags_charter <- sum(tags[, col] & school_categories == "Public charter school")
  n_tags_independent <- sum(tags[, col] & school_categories == "Independent")
  
  # Create a new row with the results and append it to table_of_differences
  new_row <- data.frame(
    Tag = tag,
    `District vs.Charter Difference` = district_vs_charter_diff,
    `Charter vs.Independent Difference` = charter_vs_independent_diff,
    `Independent vs.District Difference` = independent_vs_district_diff,
    `Pct.Tags Within(District)` = district_pct_within,
    `Pct.Tags Within(Charter)` = charter_pct_within,
    `Pct.Tags Within(Independent)` = independent_pct_within,
    `Pct.Tags Across(District)` = district_pct_across,
    `Pct.Tags Across(Charter)` = charter_pct_across,
    `Pct.Tags Across(Independent)` = independent_pct_across,
    `N Tags(District)` = n_tags_district,
    `N Tags(Charter)` = n_tags_charter,
    `N Tags(Independent)` = n_tags_independent
  )
  
  table_of_differences <- rbind(table_of_differences, new_row)
}

# Show the resulting table of differences
datatable(table_of_differences, options = list(scrollY = "400px", paging = FALSE))

```


## Top 10 Differences in charter and independent schools 
```{r charter and indep, echo=FALSE}

library(plotly)

# Filter the data for relevant columns and select the top 10 tags with the largest differences
char_ind_diff <- table_of_differences %>%
  select(Tag, "Charter.vs.Independent.Difference", "Pct.Tags.Within.Charter.", "Pct.Tags.Within.Independent.") %>%
  arrange(desc(Charter.vs.Independent.Difference)) %>%
  slice_head(n = 10)

# Create the side-by-side bar chart using plotly
char_ind_viz <- char_ind_diff %>%
  plot_ly(x = ~Tag, type = "bar", name = "Charter", y = ~`Pct.Tags.Within.Charter.`, marker = list(color = "steelblue")) %>%
  add_trace(y = ~`Pct.Tags.Within.Independent.`, name = "Independent", marker = list(color = "orange")) %>%
  layout(
    barmode = "group",
    xaxis = list(title = ""),
    yaxis = list(title = "Percent of Schools Selecting Tag"),
    title = "Top 10 Tags with Largest Differences between Charter and Independent Schools"
  )

# Display the side-by-side bar chart
char_ind_viz




```

## Top 10 Differences in Independent Vs District 
```{r sidebyside, echo=FALSE}
library(plotly)

# Filter the data for relevant columns and select the top 10 tags with the largest differences
ind_dist_diff <- table_of_differences %>%
  select(Tag, "Independent.vs.District.Difference", "Pct.Tags.Within.Independent.", "Pct.Tags.Within.District.") %>%
  arrange(desc(Independent.vs.District.Difference)) %>%
  slice_head(n = 10)

# Create the side-by-side bar chart using plotly
ind_dist_viz <- ind_dist_diff %>%
  plot_ly(x = ~Tag, type = "bar", name = "Independent", y = ~`Pct.Tags.Within.Independent.`, marker = list(color = "orange")) %>%
  add_trace(y = ~`Pct.Tags.Within.District.`, name = "District", marker = list(color = "steelblue")) %>%
  layout(
    barmode = "group",
    xaxis = list(title = ""),
    yaxis = list(title = "Percentage of Schools Selecting Tag"),
    title = "Top 10 Tags with Largest Differences between Independent and District Schools"
  )

# Display the side-by-side bar chart
ind_dist_viz


```

##Top 10 difference District and Charter
```{r moresidebyside, echo=FALSE}
library(plotly)

# Filter the data for relevant columns and select the top 10 tags with the largest differences
dist_char_diff <- table_of_differences %>%
  select(Tag, "District.vs.Charter.Difference", "Pct.Tags.Within.District.", "Pct.Tags.Within.Charter.") %>%
  arrange(desc(District.vs.Charter.Difference)) %>%
  slice_head(n = 10)

# Create the side-by-side bar chart using plotly
dist_char_viz <- dist_char_diff %>%
  plot_ly(x = ~Tag, type = "bar", name = "District", y = ~`Pct.Tags.Within.District.`, marker = list(color = "steelblue")) %>%
  add_trace(y = ~`Pct.Tags.Within.Charter.`, name = "Charter", marker = list(color = "orange")) %>%
  layout(
    barmode = "group",
    xaxis = list(title = ""),
    yaxis = list(title = "Percent of Schools Selecting Tag"),
    title = "Top 10 Tags with Largest Differences between District and Charter Schools"
  )

# Display the side-by-side bar chart
dist_char_viz


```


#Variance and variability
```{r variance cal, echo=FALSE}
# Data prep
var_all <- table_of_differences %>% 
  select(Tag, starts_with("Pct.Tags.Within.")) %>% 
  pivot_longer(cols = c("Pct.Tags.Within.Charter.", "Pct.Tags.Within.Independent.", "Pct.Tags.Within.District."),
               names_to = "type",
               names_prefix = "Pct.Tags.Within.",
               values_to = "proportion") %>% 
  ungroup() %>% 
  group_by(Tag) %>% 
  mutate(variance = sd(proportion))

# Max variability plot - top 10
var_max_10 <- var_all %>% 
  group_by(type) %>% 
  slice_max(variance, n = 10, with_ties = FALSE) %>% 
  ggplot(aes(Tag, proportion, fill = type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Tags with Most Variability",
       title = "Top 10 Tags with Largest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of increasing variability")

var_max_10

```

##Low 10 with largest variability 
```{r low var, echo=FALSE}


# Max Variability Plot - Lowest 10
var_min_10 <- var_all %>% 
  group_by(type) %>% 
  slice_min(variance, n = 10, with_ties = FALSE) %>% 
  ggplot(aes(Tag, proportion, fill = type)) + 
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Tags with Least Variability",
       title = "Top 10 Tags with Lowest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of decreasing variability")

var_min_10


```

##Top 15 with largest variability 
```{r top15, echo=FALSE}

# Max variability plot - top 15
var_max_15 <- var_all %>%
  group_by(type) %>%
  slice_max(variance, n = 15, with_ties = FALSE) %>%
  arrange(variance) %>%
  ggplot(aes(Tag, proportion, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#EF464B", "#5F9ED1", "#FBB040"),
                    labels = c("Charter", "District", "Independent"),
                    guide = guide_legend(title = "Type of School")) +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "", y = "Tags with Most Variability",
       title = "Top 15 Tags with Largest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of increasing variability")

var_max_15


```


##Lowest 15 with largest variability for charter, district, and independent
```{r low15, echo=FALSE}
# Min variability plot - lowest 15
var_min_15 <- var_all %>%
  group_by(type) %>%
  slice_min(variance, n = 15, with_ties = FALSE) %>%
  arrange(variance) %>%
  ggplot(aes(Tag, proportion, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#EF464B", "#5F9ED1", "#FBB040"),
                    labels = c("Charter", "District", "Independent"),
                    guide = guide_legend(title = "Type of School")) +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "", y = "Tags with Least Variability",
       title = "Top 15 Tags with Lowest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of increasing variability")

var_min_15

```

##Top 20 tags with Largest variability 
```{r top20variability, echo=FALSE}
# Max variability plot - top 20
var_max_20 <- var_all %>%
  group_by(type) %>%
  slice_max(variance, n = 20, with_ties = FALSE) %>%
  arrange(variance) %>%
  ggplot(aes(Tag, proportion, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#EF464B", "#5F9ED1", "#FBB040"),
                    labels = c("Charter", "District", "Independent"),
                    guide = guide_legend(title = "Type of School")) +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "", y = "Tags with Most Variability",
       title = "Top 20 Tags with Largest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of increasing variability")

var_max_20


```

##Lowest 20 tags with largest variability 

```{r}
# Min variability plot - lowest 20
var_min_20 <- var_all %>%
  group_by(type) %>%
  slice_min(variance, n = 20, with_ties = FALSE) %>%
  arrange(variance) %>%
  ggplot(aes(Tag, proportion, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#EF464B", "#5F9ED1", "#FBB040"),
                    labels = c("Charter", "District", "Independent"),
                    guide = guide_legend(title = "Type of School")) +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "", y = "Tags with Least Variability",
       title = "Top 20 Tags with Lowest Variability for Charter, District, and Independent Schools",
       subtitle = "In order of increasing variability")

var_min_20

```



